---
title: "SNAFU"
author: "Emily Pollock"
date: "1/23/2020"
output:
    html_document:
        toc: true
        toc_depth: 3
        toc_float: true
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ergm.ego)
library(EpiModel)
library(here)
source("~/Documents/Dissertation/R/Duration/Functions/netest_conversion.R")
source("~/Documents/Dissertation/R/Duration/Functions/finalEgodataPrep.R")
setwd("~/Documents/Dissertation/R/Duration/Setup/troubleshooting")
```

# Starting from scratch to troubleshoot network 
```{r}
sessionInfo()
```


## Dissolution Coefs Setup
```{r diss-coefs}
dissolution <- ~offset(edges)
exitRate <- 0.000641 # 1/(30*52) rounded
duration.marcoh <- 8.914 * 52
duration.other <- 1.824 * 52
dissolution_marcoh <- dissolution_coefs(dissolution = dissolution, duration = duration.marcoh, d.rate = exitRate)
dissolution_other <- dissolution_coefs(dissolution = dissolution, duration = duration.other, d.rate = exitRate)
```

## Edges Only - Marcoh  

sidebar: the function "ee.netest" that JKB developed has an error that only gets thrown when you use it on an edges-only model (the coef.hist doesn't get copied over correctly). I did some testing though and this doesn't seem to affect the netdx results.  

testing a few things in the following netdx runs: first a network the with n=# of nsfg respondents, with ppopsize 10,000, and then testing several different exit rates  

### ppop size = # respondents 
```{r eo-estimation, warning = F, comment = F, results='hide', fig.keep='all', message=F}
eom <- ergm.ego(egodat_marcoh ~ edges)

eom 
mcmc.diagnostics(eom)
plot(gof(eom))

#mNetest <- ee.netest(eom, dissolution_marcoh) # doesn't work b/c of coef.hist error 
mNetest2 <- ee.netest.emily(eom, dissolution_marcoh)

dx.m.static2 <- netdx(mNetest2, nsims = 1e4, dynamic = FALSE,
                    nwstats.formula = ~edges)
dx.m.static2
plot(dx.m.static2, method = "b")


#dx.m.dynamic2 <- netdx(mNetest2, nsims = 5, nsteps = 10000,
#                     nwstats.formula = ~edges)
#saveRDS(dx.m.dynamic2, here("Setup", "troubleshooting", "objects", "dx-m-dynamic.rds"))

dx.m.dynamic2 <- readRDS(here("Setup", "troubleshooting", "objects", "dx-m-dynamic.rds"))

dx.m.dynamic2
plot(dx.m.dynamic2)
plot(dx.m.dynamic2, type = "duration")
```

### ppop size = 10000
```{r eo-estimation2, warning = F, comment = F, results='hide', fig.keep='all', message=F}
eom10 <- ergm.ego(egodat_marcoh ~ edges, control = control.ergm.ego(ppopsize = 10000))

eom10
mcmc.diagnostics(eom10)
plot(gof(eom10))

mNetest10 <- ee.netest.emily(eom10, dissolution_marcoh)

dx.m.static10 <- netdx(mNetest10, nsims = 1e4, dynamic = FALSE,
                    nwstats.formula = ~edges)
dx.m.static10
plot(dx.m.static10, method = "b")

#dx.m.dynamic10 <- netdx(mNetest10, nsims = 5, nsteps = 5000,
#                     nwstats.formula = ~edges, 
#                     set.control.ergm = control.simulate.ergm(MCMC.burnin=1e6, 
#						            MCMC.interval=1e5))

#saveRDS(dx.m.dynamic10, here("Setup", "troubleshooting", "objects", "dx-m-dynamic10.rds"))

dx.m.dynamic10 <- readRDS(here("Setup", "troubleshooting", "objects", "dx-m-dynamic10.rds"))
dx.m.dynamic10
plot(dx.m.dynamic10)
plot(dx.m.dynamic10, type = "duration")
```

### higher exit rate (aging out + mortality)  
ppopsize = 10000  
```{r high-exit}
exit.high <- round(exitRate + 0.00667/52,5) # mean of yearly vital statistics (used in simulation parameters)
exit.high
dissolution_marcoh_high <- dissolution_coefs(dissolution = dissolution, d.rate = exit.high, duration = duration.marcoh)

Mnetest_high <- ee.netest.emily(eom10, dissolution_marcoh_high)

dx.m.static10high <- netdx(Mnetest_high, nsims = 1e4, dynamic = FALSE, nwstats.formula = ~edges)
dx.m.static10high
plot(dx.m.static10high)
plot(dx.m.static10high, method = "b")

#dx.m.dynamic10high <- netdx(Mnetest_high, nsims = 5, nsteps = 5000,
#                      nwstats.formula = ~edges, 
#                      set.control.ergm = control.simulate.ergm(MCMC.burnin=1e6, 
#						          MCMC.interval=1e5))

#saveRDS(dx.m.dynamic10high, here("Setup", "troubleshooting", "objects", "dx-m-dynamic10high.rds"))
dx.m.dynamic10high <- readRDS(here("Setup", "troubleshooting", "objects", "dx-m-dynamic10high.rds"))
dx.m.dynamic10high
plot(dx.m.dynamic10high)
plot(dx.m.dynamic10high, type = "duration")
```

### no exit  

```{r no-exit, message=F}
dissolution_marcoh_noexit <- dissolution_coefs(dissolution = dissolution, d.rate = 0, duration = duration.marcoh)

Mnetest_noexit <- ee.netest.emily(eom10, dissolution_marcoh_noexit)

dx.m.static10noexit <- netdx(Mnetest_noexit, nsims = 1e4, dynamic = FALSE, nwstats.formula = ~edges)
dx.m.static10noexit
plot(dx.m.static10noexit)
plot(dx.m.static10noexit, method = "b")

#dx.m.dynamic10noexit <- netdx(Mnetest_noexit, nsims = 5, nsteps = 5000,
#                      nwstats.formula = ~edges, 
#                      set.control.ergm = control.simulate.ergm(MCMC.burnin=1e6, #
#						          MCMC.interval=1e5))

#saveRDS(dx.m.dynamic10noexit, here("Setup", "troubleshooting", "objects", "dx-m-dynamic10noexit.rds"))
dx.m.dynamic10noexit <- readRDS(here("Setup", "troubleshooting", "objects", "dx-m-dynamic10noexit.rds"))
dx.m.dynamic10noexit
plot(dx.m.dynamic10noexit)
plot(dx.m.dynamic10noexit, type = "duration")
```
