---
title: "SNAFU"
author: "Emily Pollock"
date: "1/23/2020"
output:
    html_document:
        toc: true
        toc_depth: 3
        toc_float: true
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ergm.ego)
library(EpiModel)
library(here)
source("~/Documents/Dissertation/R/Duration/Functions/newnetest.R") # convert ee object to netest 
source("~/Documents/Dissertation/R/Duration/Functions/finalEgodataPrep.R") # generates the egodata objects
setwd("~/Documents/Dissertation/R/Duration/Setup/troubleshooting")
```

# Starting from scratch to troubleshoot network 
```{r}
sessionInfo()
```

# Pop Size
this is an ongoing issue, for now using network size = sample size and ppop.wt = "sample"  
pavel has some suggestions that I need to re-read and try to understand & implement.  

ppize for $3∗sampleSize/smallestWeight$  
```{r popsize, echo=FALSE}
sampSize <- nrow(egodat_marcoh$egos)
minWeight <- min(egodat_marcoh$egos$weight)

ppopsize <- 3 * (sampSize/minWeight)
```

# Age Range  
men and women aged 15:45  

# Dissolution Coefs Setup
```{r diss-coefs}
time.step <- 7  
dissolution <- ~offset(edges)

exitRate <- 0.000641 # 1/(30*52) rounded THIS MAY BE WRONG

duration.marcoh <- 8.914 * (365/time.step)
duration.other <- 1.824 * (365/time.step)

dissolution_marcoh <- dissolution_coefs(
                        dissolution = dissolution, 
                        duration = duration.marcoh, 
                        d.rate = exitRate)

dissolution_other <- dissolution_coefs(
                        dissolution = dissolution, 
                        duration = duration.other, 
                        d.rate = exitRate)
```

# Using Ergm.Ego: Edges Only - Marcoh  

sidebar: the function "ee.netest" that JKB developed has an error that only gets thrown when you use it on an edges-only model (the coef.hist doesn't get copied over correctly). I did some testing though and this doesn't seem to affect the netdx results.  

testing a few things in the following netdx runs: first a network the with n=# of nsfg respondents, and then testing several different exit rates. 

## w/ Original Exit Rate 
```{r eo-estimation, warning = F, comment = F, results='hide', fig.keep='all', message=F}
# estimate ergm.ego object 
eeo <- ergm.ego(egodat_marcoh ~ edges, control = control.ergm.ego(ppop.wt = "sample"))
# summary
summary(eeo)
# mcmc diagnostics
mcmc.diagnostics(eeo)
#gof 
plot(gof(eeo))

# convert object to netest object
est <- new.netest(eeo, dissolution_marcoh)
est

# static diagnostic
dx.static <- netdx(est, nsims = 1e5, dynamic = FALSE)
dx.static
plot(dx.static)

# dynamic diagnostic (quick)

#dx.dynamic <- netdx(est, nsims = 1, nsteps = 1e4, dynamic = T, set.control.ergm = control.simulate.ergm(MCMC.burnin = 1e5, MCMC.interval = 1e5))
#saveRDS(dx.dynamic, here("Setup", "troubleshooting", "objects", "dx-m-dynamic.rds"))

dx.dynamic <- readRDS(here("Setup", "troubleshooting", "objects", "dx-m-dynamic.rds"))

dx.dynamic
plot(dx.dynamic)
plot(dx.dynamic, type = "duration")
```

## w/ even higher exit rate (aging out + mortality)  
```{r high-exit}
# new exit rate
exit.high <- round(exitRate + 0.00667/52,5) # mean of yearly vital statistics (used in simulation parameters)
exit.high
dissolution_marcoh_high <- dissolution_coefs(dissolution = dissolution, d.rate = exit.high, duration = duration.marcoh)

# convert to netest
est_high <- new.netest(eeo, dissolution_marcoh_high)

# static diagnostic
dx.static.high <- netdx(est_high, nsims = 1e4, dynamic = FALSE)
dx.static.high
plot(dx.static.high)

#dynamic diagnostic (quick)
# dx.dynamic.high <- netdx(est_high, nsims = 1, nsteps = 5000, set.control.ergm = control.simulate.ergm(MCMC.burnin=1e5, MCMC.interval=1e5))
# saveRDS(dx.dynamic.high, here("Setup", "troubleshooting", "objects", "dx.dynamic.high.rds"))
dx.dynamic.high <- readRDS(here("Setup", "troubleshooting", "objects", "dx.dynamic.high.rds"))

dx.dynamic.high
plot(dx.dynamic.high)
plot(dx.dynamic.high, type = "duration")
```

## no exit rate 

```{r no-exit, message=F}
# dissolution params
dissolution_marcoh_noexit <- dissolution_coefs(dissolution = dissolution, d.rate = 0, duration = duration.marcoh)
# convert to netest
est_noexit <- new.netest(eeo, dissolution_marcoh_noexit)

# static diagnostics
dx.static.noexit <- netdx(est_noexit, nsims = 1e4, dynamic = FALSE)
dx.static.noexit 
plot(dx.static.noexit)

# dynamic diagnostics (quick)
#dx.dynamic.noexit <- netdx(est_noexit, nsims = 1, nsteps = 5000, set.control.ergm = control.simulate.ergm(MCMC.burnin=1e5, MCMC.interval=1e5))
#saveRDS(dx.dynamic.noexit, here("Setup", "troubleshooting", "objects", "dx.dynamic.noexit.rds"))

dx.dynamic.noexit <- readRDS(here("Setup", "troubleshooting", "objects", "dx.dynamic.noexit.rds"))
dx.dynamic.noexit
plot(dx.dynamic.noexit)
plot(dx.dynamic.noexit, type = "duration")
```

# Target Stats Approach - Marcoh  
With Original Exit Rate - this works, when this dissolution w/ ergm.ego approach loses edges quickly  
```{r targets, message=F}
# nsfg sample size
networksize <- eeo$ppopsize

# initialize same size network
net <- network.initialize(networksize, directed = FALSE)

# target stat for edges
edges.target <- est$target.stats

# dissolution - use dissolution_marcoh

# netest
fit <- netest(net, 
               formation = ~edges, 
               target.stats = edges.target,
               coef.diss = dissolution_marcoh)
summary(fit)

# static diagnostic
target.static <- netdx(fit, nsims=1000, dynamic=F)
target.static
plot(target.static)

# dynamic diagnostic
#target.dynamic <- netdx(fit, nsteps=10000, nsims=5)
#saveRDS(target.dynamic, here("Setup", "troubleshooting", "objects", "target-dynamic.rds"))

target.dynamic <- readRDS(here("Setup", "troubleshooting", "objects", "target-dynamic.rds"))
target.dynamic
plot(target.dynamic)
plot(target.dynamic, type = "duration")

```