---
title: "ergm.ego issue"
author: "Emily Pollock"
date: "11/14/2019"
output: html_document
---

```{r setup, include=FALSE}
library(ergm.ego)
library(tidyverse)
```

```{r session}
sessionInfo()
```

```{r example, error=TRUE}
# data
dat <- readRDS("~/NSFG_DATA/Objects/fullEgodata.rds")

#all alters
egodat <- egodata(egos=dat$egos, alters=dat$altersAllActive, egoWt = dat$egos$weight, egoIDcol = "ego")

# marriage/cohab alters
egodat_marcoh <- egodata(egos=dat$egos, alters=dat$altersMarCoh, egoWt = dat$egos$weight, egoIDcol = "ego")

egodat_other <- egodata(egos=dat$egos, alters=dat$altersOther, egoWt = dat$egos$weight, egoIDcol = "ego")
```

# ergm.ego issue with attribute that only exists for ego 

want: formation in "marriage/cohab" network influenced by degree in "other" network  

main problem: get odd error   

solved!   

solution: only keep deg.other.0 and deg.other.1  

why? there are so many far fewer egos with degree 2 or 3 in "other" category that ergm throws an error  

```{r counts}
egodat_marcoh$egos %>% group_by(deg.other) %>% count()
```

__version 1, where we use the default and nodefactor only drops the first category:__

```{r v1, error=TRUE}
fit.marcoh.nodecov <- ergm.ego(egodat_marcoh ~ edges + 
                                 nodecov("age") + 
                                 absdiff("sqrtage") + 
                                 nodefactor("deg.other") +
                                 offset(nodematch("sex", diff = FALSE)) +
                                 offset("concurrent"), 
                                offset.coef = c(-Inf, -Inf),
                                control = control.ergm.ego(ppopsize = 5000))
```

__version 2, where it runs but ultimately estimation fails__  

(my guess is that so few nodes who have marraiges / cohabs also have other relationships that we don't actually need this term)

```{r v2, error=TRUE}
fit.marcoh.nodecov2 <- ergm.ego(egodat_marcoh ~ edges + 
                                 nodecov("age") + 
                                 absdiff("sqrtage") + 
                                 nodefactor("deg.other", levels=c(1:2)) +
                                 offset(nodematch("sex", diff = FALSE)) +
                                 offset("concurrent"), 
                                offset.coef = c(-Inf, -Inf),
                                control = control.ergm.ego(ppopsize = 5000))
```

__"other" network example__  

```{r counts3}
egodat_marcoh$egos %>% group_by(deg.marcoh) %>% count()
```
  
version 1
```{r v-other, error=TRUE}
fit.other.cov <- ergm.ego(egodat_other ~ edges + 
                        nodecov("age") + 
                        absdiff("sqrtage") + 
                        concurrent +
                        nodefactor("deg.marcoh") +
                        offset(nodematch("sex", diff = FALSE)),
                        offset.coef = c(-Inf),
                        control = control.ergm.ego(ppopsize = 10000))
```
  
  
version 2
```{r v2-other, error=TRUE}
fit.other.cov <- ergm.ego(egodat_other ~ edges + 
                        nodecov("age") + 
                        absdiff("sqrtage") + 
                        concurrent +
                        nodefactor("deg.marcoh", levels=c(1:2)) +
                        offset(nodematch("sex", diff = FALSE)),
                        offset.coef = c(-Inf),
                        control = control.ergm.ego(ppopsize = 10000))
```