---
title: "Analysis for Network Terms"
author: "Emily Pollock"
date: "11/5/2019"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ergm.ego)
library(tidyverse)
library(DT)
library(ggplot2)
library(plotly)
library(srvyr)
library(reshape2)

dat <- readRDS("~/NSFG_DATA/Objects/fullEgodata.rds")

egodat_marcoh <- egodata(egos=dat$egos, alters=dat$altersMarCoh, egoWt = dat$egos$weight, egoIDcol = "ego")

egodat_other <- egodata(egos=dat$egos, alters=dat$altersOther, egoWt = dat$egos$weight, egoIDcol = "ego")
```

deciding which ergm terms to use based on the data

# Marriage / Cohab Network 

## Propensity to form rels by age-related term  {.tabset .tabset-pills .tabset-fade}

__couple options here:__  
  
nodefactor(age) - will give you one term for each age, not a good idea over such large age group  
  
nodefactor(agecat) = one term for each agecat - better, but really only good if each age cat has specific rates and not a general trend (if so see below)  
  
nodecov(age or agecat) - one term, assumes some monotonic rate of change between ages or agecats  

strike nodefactor(age), too many terms

let's look at new activity in last year by age and agecat

### AgeCat

```{r new-rels, echo=FALSE}
newPartsWeighted <- function(dat, range) {
  egos <- egodat_marcoh$egos$ego[which(egodat_marcoh$egos$agecat %in% range)]
  alters <- egodat_marcoh$alters[which(egodat_marcoh$alters$ego %in% egos),]
  
  svy <- as_survey(alters, weights=weight, id=1)
  
  prop <- svy %>% 
    mutate(yr = as.factor(ifelse(dfs <=12, 1, 0))) %>%
    group_by(yr) %>% 
    summarize(prop=survey_mean()) %>%
    filter(yr %in% 1) %>%
    select(-yr, -prop_se)
  
  return(prop)

}

partsW <- cbind(
                newPartsWeighted(egodat_marcoh, "15-19"), 
                newPartsWeighted(egodat_marcoh, "20-24"), 
                newPartsWeighted(egodat_marcoh, "25-29"), 
                newPartsWeighted(egodat_marcoh, "30-34"), 
                newPartsWeighted(egodat_marcoh, "35-39"), 
                newPartsWeighted(egodat_marcoh, "40-44"))

colnames(partsW) <- c("15-19","20-24","25-29", "30-34", "35-39", "40-44")
rownames(partsW) <- "Prop"
barplot(as.matrix(partsW))
datatable(round(partsW,3))

```

### Age

```{r new-rels2, echo=FALSE}
newParts2 <- function(dat, range){
  
  prop <- rep(0, length(range))
  val <- 1
  
  for (i in range) {
  egos <- dat$egos$ego[which(dat$egos$age == i)]
  alters <- dat$alters[which(dat$alters$ego %in% egos),]
  prop[val] <- nrow(alters[alters$dfs <= 12,])/nrow(alters)
  val <- val +1
  }
  return(t(round(prop,3)))
}

parts2 <- newParts2(egodat_marcoh, 15:44)
colnames(parts2) <- 15:44

barplot(parts2)
datatable(parts2)


partsByAge <- function(dat, range){
  
  l <- rep(0, length(range))
  val <- 1
  
  for (i in range) {
  egos <- egodat_marcoh$egos$ego[which(egodat_marcoh$egos$age == i)]
  alters <- egodat_marcoh$alters[which(egodat_marcoh$alters$ego %in% egos),]
  
  svy <- as_survey(alters, weights=weight, id=1)
  
  prop <- svy %>% 
    mutate(yr = as.factor(ifelse(dfs <=12, 1, 0))) %>%
    group_by(yr) %>% 
    summarize(prop=survey_mean()) %>%
    filter(yr %in% 1) %>%
    select(-yr, -prop_se)
  
  l[val] <- prop[[1]]
  
  val <- val +1
  }
  
  return(t(round(l,3)))
} 


parts3 <- partsByAge(egodat_marcoh, 15:44)
colnames(parts3) <- 15:44

barplot(parts3)
datatable(parts3)

```


## Age Mixing {.tabset .tabset-pills .tabset-fade} 
__options:__

nodematch(agecat)
absdiff / absdiffby (age / sqrt age?)

### AgeCat


```{r nodematch-marcoh, echo=FALSE}
agemixF <- round(mixingmatrix(egodat_marcoh[egodat_marcoh$egos$sex %in% "F"], "agecat", rowprob = T), 2)
amF <- melt(agemixF)
amF$sex <- "F"

agemixM <- round(mixingmatrix(egodat_marcoh[egodat_marcoh$egos$sex %in% "M"], "agecat", rowprob = T), 2)
amM <- melt(agemixM)
amM$sex <- "M"

diags <- rbind(diag(agemixF), diag(agemixM))
rownames(diags) <- c("F", "M")

md <- melt(diags)
colnames(md) <- c("sex", "agecat", "mix")

diags_p <- ggplot(md, aes(agecat, mix, fill=sex)) +
            geom_col() +
            facet_wrap(~sex) +
            ggtitle("Proportion of Within-Age-Group Ties")

ggplotly(diags_p)

datatable(diags, caption = "Within-Group Proportion of Ties")

#am <-  rbind(amF, amM)

#am %>% ggplot(aes(ego, alter)) + 
#  geom_point(color="blue", alpha=0.2, aes(size=value)) +
#  scale_size_area(max_size = 30) +
#  geom_text(aes(label=round(value,2), size=0.015)) +
#  theme(legend.position="none",
#        axis.text.x = element_text(angle=45)) +
#  coord_flip() +
#  labs(title = "Age Mixing - All Active Mar/Cohs") +
#  facet_wrap(~sex, ncol = 2)
```

### Age  


```{r agemixing, echo=FALSE}
egos <- egodat_marcoh$egos[,c("ego", "sex", "age", "sqrtage", "agesquared")]
colnames(egos) <- c("ego", "e.sex", "e.age", "e.sqrtage", "e.agesquared")
alters <- egodat_marcoh$alters[egodat_marcoh$alters$ego %in% egos$ego, 
                               c("ego", "age", "sqrtage", "agesquared", "weight")]

marcohs <- merge(alters, egos, by="ego")
marcohs <- marcohs %>% mutate(diffsqrt = e.sqrtage-sqrtage,
                              diffage = e.age-age,
                              diffsquared = e.agesquared-agesquared)

p1 <- ggplot(marcohs, aes(e.age,diffage, fill=e.sex)) +
  geom_point() +
  facet_wrap("e.sex") +
  ggtitle("Diff in Age by Ego Age")

ggplotly(p1)
```

### Sqrt Age 

```{r sqrt-age, echo=FALSE}
p2 <- ggplot(marcohs, aes(e.sqrtage, diffsqrt, fill=e.sex)) +
  geom_point() +
  facet_wrap("e.sex") +
  ggtitle("Diff in SqrtAge by Sqrt Ego Age")

ggplotly(p2)
```

### Age Squared 

```{r age-squared, echo=FALSE}
p3 <- ggplot(marcohs, aes(e.agesquared, diffsquared, fill=e.sex)) +
  geom_point() +
  facet_wrap("e.sex") +
  ggtitle("Diff in Age Squared by Ego Age Squared")

ggplotly(p3)
```

## Existing Relationships

nodefactor(deg.other, base=?)

```{r degdist, echo=FALSE}
egos <- egodat_marcoh$egos

egoSvy <- as_survey(egos, weights=weight, id=ego)

dist <- egoSvy %>% 
          filter(deg.marcoh < 2) %>%
          mutate(deg.marcoh = as.factor(deg.marcoh),
                 deg.other = as.factor(deg.other)) %>%
          group_by(deg.marcoh, deg.other) %>% 
          summarize(prop = survey_mean()) %>%
          mutate(prop = round(prop,3)) %>%
          select(-prop_se) %>%
          spread(deg.other, prop)

dist <- dist[,2:5]
colnames(dist) <- c("Deg.Other.0", "Deg.Other.1", "Deg.Other.2", "Deg.Other.3")
rownames(dist) <- c("Deg.Marcoh.0", "Deg.Marcoh.1")
datatable(dist)
```

## Offsets

sex, concurrent


<br>
# Other Network

## Propensity to form rels by age-related term  {.tabset .tabset-pills .tabset-fade}

__couple options here:__  
  
nodefactor(age) - will give you one term for each age, not a good idea over such large age group  
  
nodefactor(agecat) = one term for each agecat - better, but really only good if each age cat has specific rates and not a general trend (if so see below)  
  
nodecov(age or agecat) - one term, assumes some monotonic rate of change between ages or agecats  

strike nodefactor(age), too many terms

let's look at new activity in last year by age and agecat

### AgeCat

```{r new-rels-other, echo=FALSE}
newPartsWeighted <- function(dat, range) {
  egos <- egodat_other$egos$ego[which(egodat_other$egos$agecat %in% range)]
  alters <- egodat_other$alters[which(egodat_other$alters$ego %in% egos),]
  
  svy <- as_survey(alters, weights=weight, id=1)
  
  prop <- svy %>% 
    mutate(yr = as.factor(ifelse(dfs <=12, 1, 0))) %>%
    group_by(yr) %>% 
    summarize(prop=survey_mean()) %>%
    filter(yr %in% 1) %>%
    select(-yr, -prop_se)
  
  return(prop)

}

partsW <- cbind(
                newPartsWeighted(egodat_marcoh, "15-19"), 
                newPartsWeighted(egodat_marcoh, "20-24"), 
                newPartsWeighted(egodat_marcoh, "25-29"), 
                newPartsWeighted(egodat_marcoh, "30-34"), 
                newPartsWeighted(egodat_marcoh, "35-39"), 
                newPartsWeighted(egodat_marcoh, "40-44"))

colnames(partsW) <- c("15-19","20-24","25-29", "30-34", "35-39", "40-44")
rownames(partsW) <- "Prop"
barplot(as.matrix(partsW))
datatable(round(partsW,3))

```

### Age

```{r new-rels2-other, echo=FALSE}
newParts2 <- function(dat, range){
  
  prop <- rep(0, length(range))
  val <- 1
  
  for (i in range) {
  egos <- dat$egos$ego[which(dat$egos$age == i)]
  alters <- dat$alters[which(dat$alters$ego %in% egos),]
  prop[val] <- nrow(alters[alters$dfs <= 12,])/nrow(alters)
  val <- val +1
  }
  return(t(round(prop,3)))
}

parts2 <- newParts2(egodat_other, 15:44)
colnames(parts2) <- 15:44

barplot(parts2)
datatable(parts2)


partsByAge <- function(dat, range){
  
  l <- rep(0, length(range))
  val <- 1
  
  for (i in range) {
  egos <- egodat_other$egos$ego[which(egodat_other$egos$age == i)]
  alters <- egodat_other$alters[which(egodat_other$alters$ego %in% egos),]
  
  svy <- as_survey(alters, weights=weight, id=1)
  
  prop <- svy %>% 
    mutate(yr = as.factor(ifelse(dfs <=12, 1, 0))) %>%
    group_by(yr) %>% 
    summarize(prop=survey_mean()) %>%
    filter(yr %in% 1) %>%
    select(-yr, -prop_se)
  
  l[val] <- prop[[1]]
  
  val <- val +1
  }
  
  return(t(round(l,3)))
} 


parts3 <- partsByAge(egodat_other, 15:44)
colnames(parts3) <- 15:44

barplot(parts3)
datatable(parts3)

```

<br>
## Age Mixing {.tabset .tabset-pills .tabset-fade} 
__options:__

nodematch(agecat)
absdiff / absdiffby (age / sqrt age?)

### AgeCat


```{r nodematch-marcoh-other, echo=FALSE}
agemixF <- round(mixingmatrix(egodat_other[egodat_other$egos$sex %in% "F"], "agecat", rowprob = T), 2)
amF <- melt(agemixF)
amF$sex <- "F"

agemixM <- round(mixingmatrix(egodat_other[egodat_other$egos$sex %in% "M"], "agecat", rowprob = T), 2)
amM <- melt(agemixM)
amM$sex <- "M"

diags <- rbind(diag(agemixF), diag(agemixM))
rownames(diags) <- c("F", "M")

md <- melt(diags)
colnames(md) <- c("sex", "agecat", "mix")

diags_p <- ggplot(md, aes(agecat, mix, fill=sex)) +
            geom_col() +
            facet_wrap(~sex) +
            ggtitle("Proportion of Within-Age-Group Ties")

ggplotly(diags_p)

datatable(diags, caption = "Within-Group Proportion of Ties")

#am <-  rbind(amF, amM)

#am %>% ggplot(aes(ego, alter)) + 
#  geom_point(color="blue", alpha=0.2, aes(size=value)) +
#  scale_size_area(max_size = 30) +
#  geom_text(aes(label=round(value,2), size=0.015)) +
#  theme(legend.position="none",
#        axis.text.x = element_text(angle=45)) +
#  coord_flip() +
#  labs(title = "Age Mixing - All Active Mar/Cohs") +
#  facet_wrap(~sex, ncol = 2)
```

### Age  


```{r agemixing-other, echo=FALSE}
egos <- egodat_other$egos[,c("ego", "sex", "age", "sqrtage", "agesquared")]
colnames(egos) <- c("ego", "e.sex", "e.age", "e.sqrtage", "e.agesquared")
alters <- egodat_other$alters[egodat_other$alters$ego %in% egos$ego, 
                               c("ego", "age", "sqrtage", "agesquared", "weight")]

others <- merge(alters, egos, by="ego")
others <- others %>% mutate(diffsqrt = e.sqrtage-sqrtage,
                              diffage = e.age-age,
                              diffsquared = e.agesquared-agesquared)

p1 <- ggplot(others, aes(e.age,diffage, fill=e.sex)) +
  geom_point() +
  facet_wrap("e.sex") +
  ggtitle("Diff in Age by Ego Age")

ggplotly(p1)
```

### Sqrt Age 

```{r sqrt-age-other, echo=FALSE}
p2 <- ggplot(others, aes(e.sqrtage, diffsqrt, fill=e.sex)) +
  geom_point() +
  facet_wrap("e.sex") +
  ggtitle("Diff in SqrtAge by Sqrt Ego Age")

ggplotly(p2)
```

### Age Squared 

```{r age-squared-other, echo=FALSE}
p3 <- ggplot(others, aes(e.agesquared, diffsquared, fill=e.sex)) +
  geom_point() +
  facet_wrap("e.sex") +
  ggtitle("Diff in Age Squared by Ego Age Squared")

ggplotly(p3)
```
<br>
## Existing Relationships

nodefactor(deg.other, base=?)

```{r degdist-other, echo=FALSE}
egos <- egodat_other$egos

egoSvy <- as_survey(egos, weights=weight, id=ego)

dist <- egoSvy %>% 
          filter(deg.marcoh < 2) %>%
          mutate(deg.marcoh = as.factor(deg.marcoh),
                 deg.other = as.factor(deg.other)) %>%
          group_by(deg.marcoh, deg.other) %>% 
          summarize(prop = survey_mean()) %>%
          mutate(prop = round(prop,3)) %>%
          select(-prop_se) %>%
          spread(deg.other, prop)

dist <- dist[,2:5]
colnames(dist) <- c("Deg.Other.0", "Deg.Other.1", "Deg.Other.2", "Deg.Other.3")
rownames(dist) <- c("Deg.Marcoh.0", "Deg.Marcoh.1")
datatable(dist)
```
# Decisions 

## Marriage/Cohab Network
1. nodecov(age) seems reasonable 
2. absdiff(sqrtage)?
3. nodefactor(deg.other, levels=-1)
4. do i need a deg(0)?
5. offsets: sex, concurrent

## Other Network
1. nodefactor(agecat, levels:c(1:2)) ? (keep 15:19 and 20:24)
2. absdiff(sqrtage)
3. nodefactor(deg.marcoh, levels=-1)
4. do i need a deg(0)?
5. offsets: sex