# Parametric Models

```{r load-univariate-models, echo=FALSE}
# load univariate models

###### exponential 
e.race <- readRDS(here("model_fits", "covs", "erace.rds")) 
e.reltype <- readRDS(here("model_fits", "covs", "ereltype.rds")) 
e.partsyr <- readRDS(here("model_fits", "covs", "epartsyr.rds"))
e.network <- readRDS(here("model_fits", "covs", "enetwork.rds"))
e.agecat <- readRDS(here("model_fits", "age", "expAgeCatCurrent.rds"))
e.deg <- readRDS(here("model_fits", "covs", "edeg.rds"))

###### weibull 
w.race <- readRDS(here("model_fits", "covs", "wrace.rds")) 
w.reltype <- readRDS(here("model_fits", "covs", "wreltype.rds")) 
w.partsyr <- readRDS(here("model_fits", "covs", "wpartsyr.rds"))
w.network <- readRDS(here("model_fits", "covs", "wnetwork.rds"))
w.agecat <- readRDS(here("model_fits", "age", "weibAgeCatCurrent.rds"))
w.deg <- readRDS(here("model_fits", "covs", "wdeg.rds"))

```

## Single Covariate 

### Age Category (Current)
```{r agecat, echo=FALSE, warning=FALSE}
s <- summary(e.agecat)
plot(km_agecat_weighted, lty=2, col = c(2:6,"orange"), lwd=2,
     ylab="S(t)", xlab="Months", 
     main="Current Age Category, Exponential Dist")
lines(s$`e.agecat=15-19`$est, type = "lines", col = 2, lwd=1.5)
lines(s$`e.agecat=20-24`$est, type = "lines", col = 3, lwd=1.5)
lines(s$`e.agecat=25-29`$est, type = "lines", col = 4, lwd=1.5)
lines(s$`e.agecat=30-34`$est, type = "lines", col = 5, lwd=1.5)
lines(s$`e.agecat=35-39`$est, type = "lines", col = 6, lwd=1.5)
lines(s$`e.agecat=40-44`$est, type = "lines", col = "orange", lwd=1.5)
legend("topright", lty=c(2,1,1,1,1,1,1,1), lwd=c(rep(2,8)), col=c(1,1,2,3,4,5,6,"orange"), 
       legend = c("K-M", "Fitted Model", "15-19", "20-24", "25-29", "30-34", "35-39", "40-44"), cex = 0.8, bty="n")

s <- summary(w.agecat)
plot(km_agecat_weighted, lty=2, col = c(2:6,"orange"), lwd=2,
     ylab="S(t)", xlab="Months", 
     main="Current Age Category, Weibull Dist")
lines(s$`e.agecat=15-19`$est, type = "lines", col = 2, lwd=1.5)
lines(s$`e.agecat=20-24`$est, type = "lines", col = 3, lwd=1.5)
lines(s$`e.agecat=25-29`$est, type = "lines", col = 4, lwd=1.5)
lines(s$`e.agecat=30-34`$est, type = "lines", col = 5, lwd=1.5)
lines(s$`e.agecat=35-39`$est, type = "lines", col = 6, lwd=1.5)
lines(s$`e.agecat=40-44`$est, type = "lines", col = "orange", lwd=1.5)
legend("topright", lty=c(2,rep(1,7)), lwd=c(rep(2,8)), col=c(1,1,2,3,4,5,6,"orange"), 
       legend = c("K-M", "Fitted Model", "15-19", "20-24", "25-29", "30-34", "35-39", "40-44"), cex = 0.8, bty="n")
```

### Race

```{r race-cov, echo=FALSE}
plot(e.race, main = "Ego Race, Exp Dist", 
     col=c("red", "blue", "springgreen3", "orange"), xlab="t, months")
legend(370, 0.95, legend = c("black", "white", "hisp", "other"), col=c("blue", "red", "springgreen3", "orange"), lwd=2, lty=1, cex=.7)
plot(w.race, main = "Ego Race, Weibull Dist", 
     col=c("red", "blue", "springgreen3", "orange"), xlab="t, months")
legend(370, 0.95, legend = c("black", "white", "hisp", "other"), col=c("blue", "red", "springgreen3", "orange"), lwd=2, lty=1, cex=.7)
```

### Rel Type
```{r reltype-cov, echo=FALSE}
plot(e.reltype, main = "Rel Type, Exp Dist", 
     col=c("red", "blue", "springgreen3"), xlab="t, months")
legend(370, 1, legend = c("Marriage", "Cohab", "Other"), col=c("red", "springgreen3", "blue"), lwd=2, lty=1, cex=.7)

plot(w.reltype, main = "Rel Type, Weibull Dist", 
     col=c("red", "blue", "springgreen3"), xlab="t, months")
legend(370, 1, legend = c("Marriage", "Cohab", "Other"), col=c("red", "springgreen3", "blue"), lwd=2, lty=1, cex=.7)
```

### Network  
Similar to above, but split into marcoh and other - two of the networks we often split out
```{r network-cov, echo=FALSE}
plot(e.network, main = "Network, Exp Dist", 
     col=c("red", "blue"), xlab="t, months")
legend(370, 1, legend = c("Mar/Coh", "Other"), col=c("red","blue"), lwd=2, lty=1, cex=.7)

plot(w.network, main = "Network, Weibull Dist", 
     col=c("red", "blue"), xlab="t, months")
legend(370, 1, legend = c("Mar/Coh", "Other"), col=c("red","blue"), lwd=2, lty=1, cex=.7)
```

### Partners in Last Year

Note: this was also run with "osnpyr" and "maxospyr" with very little differences (not shown)  

```{r partsyr-cov, echo=FALSE}
plot(e.partsyr, main = "Partners in Last Year, Exp Dist", 
     col=c("red", "blue", "springgreen3"), xlab="t, months")
legend(370, 1, legend = c("1", "2", "3+"), col=c("red", "springgreen3", "blue"), lwd=2, lty=1, cex=.7)

plot(w.partsyr, main = "Partners in Last Year, Weibull Dist", 
     col=c("red", "blue", "springgreen3"), xlab="t, months")
legend(370, 1, legend = c("1", "2", "3+"), col=c("red", "springgreen3", "blue"), lwd=2, lty=1, cex=.7)
```

### Momentary Degree
```{r deg-cov, echo=FALSE}
plot(e.deg, main = "Momentary Degree, Exp Dist",
     col=2:5, xlab="t, months")
legend("right", 1, legend = c("0", "1", "2", "3"), col=c(3,2,5,4), lwd=2, lty=1, cex=.7)

plot(w.deg, main = "Momentary Degree, Weibull Dist",
     col=2:5, xlab="t, months")
legend("right", 1, legend = c("0", "1", "2", "3"), col=c(3,2,5,4), lwd=2, lty=1, cex=.7)
```


### AIC Comparison 

Model | Exp Dist | Weibull 
------|----------|--------
No Covs | `r AIC(expFit)` | `r AIC(weibFit)`
Age Category |  `r AIC(e.agecat)` |  `r AIC(w.agecat)`
Race | `r AIC(e.race)` | `r AIC(w.race)`
Rel Type - 3 |  `r AIC(e.reltype)` | `r AIC(w.reltype)`
Rel Type - 2 | `r AIC(e.network)` | `r AIC(w.network)`
Parts in Last Year | `r AIC(e.partsyr)` | `r AIC(w.partsyr)`
Momentary Degree | `r AIC(e.deg)` | `r AIC(w.deg)` 



## Multiple 

these plots are messy 

```{r load-multivariate-models, echo=FALSE}
# load multivariate models

###### exponential 
e.raceage <- readRDS(here("model_fits", "covs", "eraceage.rds")) 

e.racerels <- readRDS(here("model_fits", "covs", "eracerels.rds"))
e.agerels <- readRDS(here("model_fits", "covs", "eagerels.rds"))

e.ageparts <- readRDS(here("model_fits", "covs", "eageparts.rds")) 
e.raceparts <- readRDS(here("model_fits", "covs", "eraceparts.rds")) 

e.agedeg <- readRDS(here("model_fits", "covs", "eagedeg.rds")) 
e.racedeg <- readRDS(here("model_fits", "covs", "eracedeg.rds"))

e.relparts <- readRDS(here("model_fits", "covs", "erelparts.rds"))

###### weibull 
w.raceage <- readRDS(here("model_fits", "covs", "wraceage.rds"))

w.racerels <- readRDS(here("model_fits", "covs", "wracerels.rds"))
w.agerels <- readRDS(here("model_fits", "covs", "wagerels.rds"))

w.ageparts <- readRDS(here("model_fits", "covs", "wageparts.rds")) 
w.raceparts <- readRDS(here("model_fits", "covs", "wraceparts.rds")) 

w.agedeg <- readRDS(here("model_fits", "covs", "wagedeg.rds")) 
w.racedeg <- readRDS(here("model_fits", "covs", "wracedeg.rds"))

w.relparts <- readRDS(here("model_fits", "covs", "wrelparts.rds"))
```

### Age Cat + Race
```{r agerace, echo=FALSE}
plot(e.agerels, main = "Age Cat + Rel Type, Exp Dist", 
     xlab="t, months")

plot(w.agerels, main = "Age Cat + Rel Type, Weibull Dist", 
     xlab="t, months")
```

### AgeCat + Rel Type
```{r agerels, echo=FALSE}
plot(e.ageparts, main = "Age Cat + Partners in Last Year, Exp Dist", 
     xlab="t, months")
#col=c("red", "blue", "springgreen3"), 
#legend(370, 1, legend = c("Marriage", "Cohab", "Other"), col=c("red", "springgreen3", "blue"), lwd=2, lty=1, cex=.7)

plot(w.ageparts, main = "Age Cat + Partners in Last Year, Weibull Dist", 
     xlab="t, months")

```

### Race + Rel Type  

```{r racerels, echo=FALSE}
plot(e.racerels, main = "Race + Rel Type, Exp Dist", 
     xlab="t, months")
     
plot(w.racerels, main = "Race + Rel Type, Weibull Dist", 
     xlab="t, months")
```

### Age Cat + Partners in Last Year

```{r ageparts, echo=FALSE}
plot(e.ageparts, main = "Age Cat + Partners in Last Year, Exp Dist", 
     xlab="t, months")
#col=c("red", "blue", "springgreen3"), 
#legend(370, 1, legend = c("Marriage", "Cohab", "Other"), col=c("red", "springgreen3", "blue"), lwd=2, lty=1, cex=.7)

plot(w.ageparts, main = "Age Cat + Partners in Last Year, Weibull Dist", 
     xlab="t, months")

```

### Race + Partners in Last Year

```{r raceparts, echo=FALSE}
plot(e.raceparts, main = "Race + Partners in Last Year, Exp Dist", 
     xlab="t, months")

plot(w.raceparts, main = "Race + Partners in Last Year, Weibull Dist", 
     xlab="t, months")
```


### Age Cat + Momentary Degree

```{r agedeg, echo=FALSE}
plot(e.agedeg, main = "Age Cat + Momentary Degree, Exp Dist", 
     xlab="t, months")

plot(w.agedeg, main = "Age Cat + Momentary Degree, Weibull Dist", 
     xlab="t, months")

```


### Race + Momentary Degree
```{r racedeg, echo=FALSE}
plot(e.racedeg, main = "Race + Momentary Degree, Exp Dist", 
     xlab="t, months")

plot(w.racedeg, main = "Race + Momentary Degree, Weibull Dist", 
     xlab="t, months")

```

### Rel Type + Parts Last Year

```{r relparts, echo=FALSE}
plot(e.relparts, main = "Rel Type + Parts Lst Year, Exp Dist", 
     xlab="t, months")

plot(w.relparts, main = "Rel Type + Parts Lst Year, Weibull Dist", 
     xlab="t, months")
```



### Model Comparison 
Model | Exp Dist | Weibull 
------|----------|--------
No Covs | `r AIC(expFit)` | `r AIC(weibFit)`
Age Cat + Race |  `r AIC(e.raceage)` |  `r AIC(w.raceage)`
Age Cat + Rel Type |  `r AIC(e.agerels)` | `r AIC(w.agerels)`
Race + Rel Type | `r AIC(e.racerels)` | `r AIC(w.racerels)`
Age Cat + PartsLstYear | `r AIC(e.ageparts)` | `r AIC(w.ageparts)`
Race + PartsLastYear | `r AIC(e.raceparts)` | `r AIC(w.raceparts)`
Age Cat + Degree | `r AIC(e.agedeg)` | `r AIC(w.agedeg)`
Race + Degree | `r AIC(e.racedeg)` | `r AIC(w.racedeg)`
Rel Type + PartsLstYear | `r AIC(e.relparts)` | `r AIC(w.relparts)`





### Other Age Vars 
 that don't tell us much 

```{r full-dataset, echo=FALSE, warnings=FALSE}
######## load continuous covariate models ##########
exp <- readRDS(here("model_fits", "nocovs", "exp.rds")) 
weib <- readRDS(here("model_fits", "nocovs", "weib.rds")) 

expEgoAgeCurrent <- readRDS(here("model_fits", "age", "expEgoAgeCurrent.rds"))
weibEgoAgeCurrent <- readRDS(here("model_fits", "age", "weibEgoAgeCurrent.rds"))

expEgoAgeInit <- readRDS(here("model_fits", "age", "expEgoAgeInit.rds"))
weibEgoAgeInit <- readRDS(here("model_fits", "age", "weibEgoAgeInit.rds"))

# diff sqrt age
expAgeDiff <- readRDS(here("model_fits", "age", "expAgeDiff.rds"))
weibAgeDiff <- readRDS(here("model_fits", "age", "weibAgeDiff.rds"))

expAgeDiff_CurAge <- readRDS(here("model_fits", "age", "expAgeDiff_CurAge.rds"))
weibAgeDiff_CurAge <- readRDS(here("model_fits", "age", "weibAgeDiff_CurAge.rds"))

expAgeDiffAgeCatCurrent <- readRDS(here("model_fits", "age", "expAgeDiffAgeCatCurrent.rds"))
weibAgeDiffAgeCatCurrent <- readRDS(here("model_fits", "age", "weibAgeDiffAgeCatCurrent.rds"))
```

```{r exp-plots, out.width='50%', echo=FALSE}

#plot non-categorical models, first one by one and then all together

plot(exp, main="Exponential, No Covs",
     ylab="S(t)", xlab="t, months", lwd=3, col="red")
legend(300, 1, legend = c("K-M", "No Covs"), 
       col = c(1, "red"), lwd=3)


plot(expEgoAgeCurrent,
     main="Exponential, Current Ego Age",
     ylab="S(t)", xlab="t, months", lwd=3, col="steelblue3")
legend(300, 1, legend = c("K-M", "Current Ego Age"), 
       col = c(1, "steelblue3"), lwd=3)

plot(expEgoAgeInit,
     main="Exponential, Initial Ego Age",
     ylab="S(t)", xlab="t, months", lwd=3, col="orange")
legend(300, 1, legend = c("K-M", "Initial Ego Age"), 
       col = c(1, "orange"), lwd=3)

plot(expAgeDiff,
     main="Exponential, Age Difference",
     ylab="S(t)", xlab="t, months", lwd=3, col="springgreen3")
legend(300, 1, legend = c("K-M", "Age Difference"), 
       col = c(1, "springgreen3"), lwd=3)
```

```{r exp-plot-full, echo=FALSE}
plot(exp,
     main="Basic Age Dynamics, Exp Dist",
     ylab="S(t)", xlab="t, months", lwd=3)
lines(expAgeDiff, col="forestgreen")
lines(expEgoAgeInit, col="orange", lwd=3)
lines(expEgoAgeCurrent, col="steelblue3", lwd=3)
legend(300, 1, legend = c("K-M", "No Covs", "AgeDiff", "Initial Ego Age", "Current Ego Age"), 
       col = c(1, "red", "springgreen3", "orange", "steelblue3"), lwd=3)
```

```{r weib-plots, echo=FALSE, out.width='50%'}
plot(weib, main="Weibull, No Covs", ylab="S(t)", xlab="t, months", lwd=3, col="red")
legend(300, 1, legend=c("KM", "No Covs"), col = c(1, "red"), lwd=3)

plot(weibAgeDiff, main="Weibull, Age Diff", ylab="S(t)", xlab="t, months", lwd=3, col="springgreen3")
legend(300, 1, legend=c("KM", "Age Diff"), col = c(1, "springgreen3"), lwd=3)

plot(weibEgoAgeInit, main="Weibull, Initial Ego Age", ylab="S(t)", xlab="t, months", lwd=3, col="orange")
legend(300, 1, legend=c("KM", "Initial Ego Age"), col = c(1, "orange"), lwd=3)

plot(weibEgoAgeCurrent, main="Weibull, Current Ego Age", ylab="S(t)", xlab="t, months", lwd=3, col="steelblue3")
legend(300, 1, legend=c("KM", "Current Ego Age"), col = c(1, "steelblue3"), lwd=3)
```

```{r weib-plot-full, echo=FALSE}
plot(weib,
     main="Basic Age Dynamics, Weibull Dist",
     ylab="S(t)", xlab="t, months", lwd=3)
lines(weibAgeDiff, col="springgreen3")
lines(weibEgoAgeInit, col="orange", lwd=3)
lines(weibEgoAgeCurrent, col="steelblue3", lwd=3)
legend(300, 1, legend = c("K-M", "No Covs", "AgeDiff", "Initial Ego Age", "Current Ego Age"), 
       col = c(1, "red", "springgreen3", "orange", "steelblue3"), lwd=3)

```


## Thoughts & Discussions 

 - age is not responsible for the initial steep drop-off
 - exp mixture models & weibulls fit better because there is failty endogenous to the relationships 
 - weibull isn't masking the effect of age
 - has anybody demonstrated a crosswalk between a exponential mixture and a weibull? 
 - theory: why doesn't age have a larger effect?
 - age / period / cohort effects - what would we expect? 
 
 
 __Age’s effects on relationship duration are complex__  
 - Relationship age at time of interview usually lower for younger people, but due to a combination of ‘churn’ and not being alive long enough to have had decades-long relationships. 
 - Age difference between ego and alter matters, but affect of age difference is likely to vary across the life course. 
 - The effect age of age difference on relationship duration may be different depending on if the male or female is the older partner. 
 - Which age is most important for relationship persistence? Age at beginning of relationship, or current age? 
 - Whose age to model? Ego age? Partner age? Ego age and age difference? 

__Dyadic Complications__  
 - Relationship duration is unlikely to be independent of other relationships each partner may be participating in. 
 - Is there a way to represent this in a parametric regression model, and if not, what are the consequences of assuming independence between relationships? 

Likely other complications. 
