# CSSS Poster

## Histograms
### Full Distribtuion, 15-44 year olds

looks exponential! so why do most of our models based on an exponential distribution fail to represent the data? look below by age cat...

```{r hist-full, echo=FALSE}
ggplot(dat, aes(x = t_c_years, fill = rel2)) +
  geom_histogram(binwidth = 0.5) + 
  scale_x_continuous(name="years", breaks=seq(0,35,5)) + 
  ggtitle("Relationship Age, Full Dataset (Ages 15-44)")
```

### By Age Category

First 3 age cats, the 15-29 year olds (& my pop of interest for modeling purposes), look reasonably exponential. But! Older age categories begin to look surprisingly uniform  


```{r histograms, echo=FALSE, out.width='50%'}
dat %>% filter(e.agecat %in% "15-19") %>%
  ggplot(aes(x = t_c_years, fill = rel2, weights = e.weight)) +
  geom_histogram(binwidth = 0.5) +
  scale_x_continuous(name="years", limits = c(0,40)) +
    coord_cartesian(ylim=c(0,1e7)) +
  scale_fill_discrete(name="Rel Type") + 
  ggtitle("Ages 15-19")

dat %>% filter(e.agecat %in% "20-24") %>%
  ggplot(aes(x = t_c_years, fill = rel2, weights = e.weight)) +
  geom_histogram(binwidth = 0.5) +
  scale_x_continuous(name="years", limits = c(0,40)) +
  coord_cartesian(ylim=c(0,1e7)) +
  scale_fill_discrete(name="Rel Type") + 
  ggtitle("Ages 20-24")

dat %>% filter(e.agecat %in% "25-29") %>%
  ggplot(aes(x = t_c_years, fill = rel2, weights = e.weight)) +
  geom_histogram(binwidth = 0.5) +
  scale_x_continuous(name="years", limits = c(0,40)) +
  coord_cartesian(ylim=c(0,1e7)) +
  scale_fill_discrete(name="Rel Type") + 
  ggtitle("Ages 25-29")


dat %>% filter(e.agecat %in% "30-34") %>%
  ggplot(aes(x = t_c_years, fill = rel2, weights = e.weight)) +
  geom_histogram(binwidth = 0.5) +
  scale_x_continuous(name="years", limits = c(0,40)) +
  coord_cartesian(ylim=c(0,1e7)) +
  scale_fill_discrete(name="Rel Type") + 
  ggtitle("Ages 30-34")

dat %>% filter(e.agecat %in% "35-39") %>%
  ggplot(aes(x = t_c_years, fill = rel2, weights = e.weight)) +
  geom_histogram(binwidth = 0.5) +
  scale_x_continuous(name="years", limits = c(0,40)) +
  coord_cartesian(ylim=c(0,1e7)) +
  scale_fill_discrete(name="Rel Type") + 
  ggtitle("Ages 35-39")

dat %>% filter(e.agecat %in% "40-44") %>%
  ggplot(aes(x = t_c_years, fill = rel2, weights = e.weight)) +
  geom_histogram(binwidth = 0.5) +
  scale_x_continuous(name="years", limits = c(0,40)) +
  coord_cartesian(ylim=c(0,1e7)) +
  scale_fill_discrete(name="Rel Type") + 
  ggtitle("Ages 40-44")

```


## test graphs

trying to lay latent models over histograms 
first make hists unweighted since latent models are unweighted 

```{r hist-mixexp-full, echo=FALSE}

ggplot(dat, aes(x = t_c_years, fill = rel2)) +
  geom_histogram(binwidth = 0.5) + 
  scale_x_continuous(name="years", breaks=seq(0,35,5)) + 
  ggtitle("Relationship Age, Full Dataset (Ages 15-44)")

durVec <- 0:max(dat$t_c)
expfull <- invLogit(coef(mixExpFit)[3]) *
      pexp(q = durVec, exp(coef(mixExpFit)[1]), lower.tail = FALSE) +
  
      (1-invLogit(coef(mixExpFit)[3])) *
      pexp(q = durVec, exp(coef(mixExpFit)[2]), lower.tail = FALSE)

mixexp <- as.data.frame(cbind(durVec, expfull))
mixexp$durVec <- mixexp$durVec/12

ggplot() +
  geom_histogram(data = dat, aes(x = t_c_years, fill = rel2), binwidth = 0.5) + 
  geom_line(data = mixexp, aes(x=durVec, y=expfull*7175)) +
  scale_x_continuous(name="years", breaks=seq(0,35,5)) +
  ggtitle("Relationship Age, Full Dataset (Ages 15-44)") 

ggplot() +
  geom_histogram(data = dat, aes(x = t_c_years, fill = rel2), binwidth = 0.5) + 
  geom_line(data = mixexp, aes(x=durVec, y=expfull*length(dat$ego))) +
  scale_x_continuous(name="years", breaks=seq(0,35,5)) +
  ggtitle("Relationship Age, Full Dataset (Ages 15-44)") 

newdat <- dat[order(dat$t_c_years),]

ggplot() +
  geom_bar(data = newdat, aes(x = t_c_years, y = ego, fill = rel2), stat="identity") + 
  #geom_line(data = mixexp, aes(x=durVec, y=expfull*length(dat$ego))) +
  scale_x_continuous(name="years", breaks=seq(0,35,5)) +
  ggtitle("Relationship Age, Full Dataset (Ages 15-44)") 

```