# Mixture Models 


```{r load-mixes, echo=FALSE, include=FALSE}
## Weighted vs Unweighted  
mixExpFitweighted <- readRDS(here("model_fits", "mixed", "mix-exp-nsfg-weighted.rds"))
#mixExpFit <- readRDS(here("model_fits", "mixed", "mix-exp-nsfg.rds"))
#km_weighted <- readRDS(here("model_fits", "km", "km-weighted.rds"))

```

```{r mix-plots, echo=FALSE, include=FALSE}
durVec <- c(1:max(km_weighted$time))

plot(km_weighted, 
     conf.int = FALSE,
     xlab = "t, months", ylab = "S(t)",
     main="Mixed Exponential, No Covs"
)
lines(x = durVec,
      y = invLogit(coef(mixExpFit)[3]) *
      pexp(q = durVec, exp(coef(mixExpFit)[1]), lower.tail = FALSE) +
      (1-invLogit(coef(mixExpFit)[3])) *
      pexp(q = durVec, exp(coef(mixExpFit)[2]), lower.tail = FALSE),
      col ="springgreen3", lwd=2
)
lines(x = durVec,
      y = invLogit(coef(mixExpFitweighted)[3]) *
      pexp(q = durVec, exp(coef(mixExpFitweighted)[1]), lower.tail = FALSE) +
      (1-invLogit(coef(mixExpFitweighted)[3])) *
      pexp(q = durVec, exp(coef(mixExpFitweighted)[2]), lower.tail = FALSE),
      col ="purple", lwd=2
)
legend(300,1, legend = c("Kaplan-Meier", "Mixed Exp (no weights)", "Mixed Exp Weighted"),
col = c("black", "springgreen3", "purple"), cex=0.7, lwd=2)
```

__Mixed Exp Models by Current AgeCat__

Mixture model across whole pop does pretty well but overestimates duration of young rels and underestmates duration of long rels. Let's see if we can do better when each age cat is its own mixture model (2 lambdas and a mixture weight). 

Here I fit 2-component mixture models to each of the age categories and plot them as separate lines. 

Two models categories are displayed - 

1) a series of mixed exponential models that were fit to specific ego age categories, which when are averaged according to pop size to compare to overall K-M and other models
(this means that each age category has its own two rate params (rel dur) and proportion of rels in each distribution (weight))

2) one mixed exponential model with two rate params that apply to all age groups, but the weight changes by age category. The first model should fit K-M better, but this is to see if one short/long duration is a reasonable simplifying assumption (with weights changing by age)



## Mixed Exp #1  

a series of mixed exponential models that were fit to specific ego age categories, which when are averaged according to pop size to compare to overall K-M and other models
(this means that each age category has its own two rate params (rel dur) and proportion of rels in each distribution (weight))


```{r mixed-exp-ea, echo=FALSE, warning=FALSE, message=FALSE}

km_agecat_weighted <- readRDS(here("model_fits",  "km", "km_ac_weighted.rds"))


##### load models #######
mix1519 <- readRDS(here("model_fits", "mixed", "mix1519weighted.rds"))
mix2024 <- readRDS(here("model_fits", "mixed", "mix2024weighted.rds"))
mix2529 <- readRDS(here("model_fits", "mixed", "mix2529weighted.rds"))
mix3034 <- readRDS(here("model_fits", "mixed", "mix3034weighted.rds"))
mix3539 <- readRDS(here("model_fits", "mixed", "mix3539weighted.rds"))
mix4044 <- readRDS(here("model_fits", "mixed", "mix4044weighted.rds"))

# inverse logit function
logit = function(p){
  log(p/(1-p))
}

invLogit = function(phi){
  1/(1+exp(-phi))
}

###### plot using plotly #######
durVec <- c(1:max(km_agecat_weighted$time))

# made data frame of mixed exp outputs
df <- as.data.frame(cbind(durVec, 
            invLogit(coef(mix1519)[3]) * pexp(q = durVec, exp(coef(mix1519)[1]), lower.tail = FALSE) +
        (1-invLogit(coef(mix1519)[3])) * pexp(q = durVec, exp(coef(mix1519)[2]), lower.tail = FALSE),
            invLogit(coef(mix2024)[3]) * pexp(q = durVec, exp(coef(mix2024)[1]), lower.tail = FALSE) +
        (1-invLogit(coef(mix2024)[3])) * pexp(q = durVec, exp(coef(mix2024)[2]), lower.tail = FALSE), 
            invLogit(coef(mix2529)[3]) * pexp(q = durVec, exp(coef(mix2529)[1]), lower.tail = FALSE) +
        (1-invLogit(coef(mix2529)[3])) * pexp(q = durVec, exp(coef(mix2529)[2]), lower.tail = FALSE),
            invLogit(coef(mix3034)[3]) * pexp(q = durVec, exp(coef(mix3034)[1]), lower.tail = FALSE) +
        (1-invLogit(coef(mix3034)[3])) * pexp(q = durVec, exp(coef(mix3034)[2]), lower.tail = FALSE),
            invLogit(coef(mix3539)[3]) * pexp(q = durVec, exp(coef(mix3539)[1]), lower.tail = FALSE) +
        (1-invLogit(coef(mix3539)[3])) * pexp(q = durVec, exp(coef(mix3539)[2]), lower.tail = FALSE),
            invLogit(coef(mix4044)[3]) * pexp(q = durVec, exp(coef(mix4044)[1]), lower.tail = FALSE) +
        (1-invLogit(coef(mix4044)[3])) * pexp(q = durVec, exp(coef(mix4044)[2]), lower.tail = FALSE)))
colnames(df) <- c("t", "15-19", "20-24", "25-29", "30-34", "35-39", "40-44")


plot(km_agecat_weighted, lty=2, col = c(2:6,"orange"), lwd=2,
     ylab="S(t)", xlab="Months", 
     main="Current Age Category, 2-Component Mixed Exponential")
lines(df$`15-19`, type = "lines", col = 2, lwd=1.5)
lines(df$`20-24`, type = "lines", col = 3, lwd=1.5)
lines(df$`25-29`, type = "lines", col = 4, lwd=1.5)
lines(df$`30-34`, type = "lines", col = 5, lwd=1.5)
lines(df$`35-39`, type = "lines", col = 6, lwd=1.5)
lines(df$`40-44`, type = "lines", col = "orange", lwd=1.5)
legend("topright", lty=c(2,1,1,1,1,1,1,1), lwd=c(rep(2,8)), col=c(1,1,2,3,4,5,6,"orange"), 
       legend = c("K-M", "Fitted Model", "15-19", "20-24", "25-29", "30-34", "35-39", "40-44"), cex = 0.8, bty="n")


# weighted average model results 
p <- dat %>% count(e.agecat) %>% mutate(prop = n/sum(n))

df$w15 <- as.numeric(rep(p[1,3], length(df$t)))
df$w20 <- as.numeric(rep(p[2,3], length(df$t)))
df$w25 <- as.numeric(rep(p[3,3], length(df$t)))
df$w30 <- as.numeric(rep(p[4,3], length(df$t)))
df$w35 <- as.numeric(rep(p[5,3], length(df$t)))
df$w40 <- as.numeric(rep(p[6,3], length(df$t)))

avg <- df$`15-19`*df$w15 + df$`20-24`*df$w20 + df$`25-29`*df$w25 + 
       df$`30-34`*df$w30 + df$`35-39`*df$w35 + df$`40-44`*df$w40

# compare avg to survival 
df2 <- as.data.frame(cbind(durVec, avg))
colnames(df2) <- c("t", "MixedExpAvg")

plot(km_weighted, main="KM vs Average of Mixed Exponential")
lines(x=df2$t, y=df2$MixedExpAvg, col="purple")


#plot_ly(df2, x = ~t/12, y = ~KM, mode = 'lines', name = "K-M, unweighted", type='scatter') %>%
#  add_trace(y = ~MixedExpAvg, mode = 'lines', name = "Mixed Exp") %>%
#  layout(title="K-M versus Weighted Average of Mixed Exp Models (no survey weights)",
#         yaxis = list(title="S(t)"), 
#         xaxis= list("t, years"))

```


```{r mixexp-table, echo=FALSE}
rate15_1 <- (1/exp(coef(mix1519)[1]))/12
rate15_2 <- (1/exp(coef(mix1519)[2]))/12
weight15 <- invLogit(coef(mix1519)[3])

rate20_1 <- (1/exp(coef(mix2024)[1]))/12
rate20_2 <- (1/exp(coef(mix2024)[2]))/12
weight20 <- invLogit(coef(mix2024)[3])

rate25_1 <- (1/exp(coef(mix2529)[1]))/12
rate25_2 <- (1/exp(coef(mix2529)[2]))/12
weight25 <- invLogit(coef(mix2529)[3])

rate30_1 <- (1/exp(coef(mix3034)[1]))/12
rate30_2 <- (1/exp(coef(mix3034)[2]))/12
weight30 <- invLogit(coef(mix3034)[3])

rate35_1 <- (1/exp(coef(mix3539)[1]))/12
rate35_2 <- (1/exp(coef(mix3539)[2]))/12
weight35 <- invLogit(coef(mix3539)[3])

rate40_1 <- (1/exp(coef(mix4044)[1]))/12
rate40_2 <- (1/exp(coef(mix4044)[2]))/12
weight40 <- invLogit(coef(mix4044)[3])
```


Age Category | Short Dur (years) | Long Dur (years) | Prop. Short  
---------- | -----------|----------|-------------
15-19 | `r round(rate15_1,2)` | `r round(rate15_2,2)` | `r round(weight15,2)`
20-24 | `r round(rate20_1,2)` | `r round(rate20_2,2)` | `r round(weight20,2)`
25-29 | `r round(rate25_1,2)` | `r round(rate25_2,2)` | `r round(weight25,2)`
30-34 | `r round(rate30_1,2)` | `r round(rate30_2,2)` | `r round(weight30,2)`
35-39 | `r round(rate35_1,2)` | `r round(rate35_2,2)` | `r round(weight35,2)`
40-44 | `r round(rate40_1,2)` | `r round(rate40_2,2)` | `r round(weight40,2)`


## Mixed Exp #2

generalization of above: one mixed exponential model with two rate params that apply to all age groups, but the weight changes by age category. The first model should fit K-M better, but this is to see if one short/long duration is a reasonable simplifying assumption (with weights changing by age)  


```{r compare-mixexps, echo=FALSE}
# compare avg to survival 
weightsMixExp <- readRDS(here("model_fits", "mixed", "weightsMixExp_surveyweighted.rds"))

shortDur <- 1/exp(coef(weightsMixExp)[1])
longDur <- 1/exp(coef(weightsMixExp)[2])
wmix15 <- invLogit(coef(weightsMixExp)[3])
wmix20 <- invLogit(coef(weightsMixExp)[4])
wmix25 <- invLogit(coef(weightsMixExp)[5])
wmix30 <- invLogit(coef(weightsMixExp)[6])
wmix35 <- invLogit(coef(weightsMixExp)[7])
wmix40 <- invLogit(coef(weightsMixExp)[8])

dfw <- as.data.frame(cbind(durVec, 
            wmix15 * pexp(q = durVec, exp(coef(weightsMixExp)[1]), lower.tail = FALSE) +
        (1-wmix15) * pexp(q = durVec, exp(coef(weightsMixExp)[2]), lower.tail = FALSE),
            wmix20 * pexp(q = durVec, exp(coef(weightsMixExp)[1]), lower.tail = FALSE) +
        (1-wmix20) * pexp(q = durVec, exp(coef(weightsMixExp)[2]), lower.tail = FALSE), 
            wmix25 * pexp(q = durVec, exp(coef(weightsMixExp)[1]), lower.tail = FALSE) +
        (1-wmix25) * pexp(q = durVec, exp(coef(weightsMixExp)[2]), lower.tail = FALSE),
            wmix30 * pexp(q = durVec, exp(coef(weightsMixExp)[1]), lower.tail = FALSE) +
        (1-wmix30) * pexp(q = durVec, exp(coef(weightsMixExp)[2]), lower.tail = FALSE),
            wmix35 * pexp(q = durVec, exp(coef(weightsMixExp)[1]), lower.tail = FALSE) +
        (1-wmix35) * pexp(q = durVec, exp(coef(weightsMixExp)[2]), lower.tail = FALSE),
            wmix40 * pexp(q = durVec, exp(coef(weightsMixExp)[1]), lower.tail = FALSE) +
        (1-wmix40) * pexp(q = durVec, exp(coef(weightsMixExp)[2]), lower.tail = FALSE)))

colnames(dfw) <- c("t", "15-19", "20-24", "25-29", "30-34", "35-39", "40-44")

# weighted average model results 
dfw$w15 <- as.numeric(rep(p[1,3], length(dfw$t)))
dfw$w20 <- as.numeric(rep(p[2,3], length(dfw$t)))
dfw$w25 <- as.numeric(rep(p[3,3], length(dfw$t)))
dfw$w30 <- as.numeric(rep(p[4,3], length(dfw$t)))
dfw$w35 <- as.numeric(rep(p[5,3], length(dfw$t)))
dfw$w40 <- as.numeric(rep(p[6,3], length(dfw$t)))

avg_w <- dfw$`15-19`*dfw$w15 + dfw$`20-24`*dfw$w20 + dfw$`25-29`*dfw$w25 + 
       dfw$`30-34`*dfw$w30 + dfw$`35-39`*dfw$w35 + dfw$`40-44`*dfw$w40

dfw2 <- as.data.frame(cbind(durVec, avg, avg_w))
colnames(dfw2) <- c("t", "MixedExpAvg", "WeightsMixExp")

plot(km_weighted, main = "KM versus AgeCat Mixed Exps Models")
lines(x=dfw2$t, y=dfw2$MixedExpAvg, col="purple", lwd=2)
lines(x=dfw2$t, y=dfw2$WeightsMixExp, col="blue", lwd=2)
legend("topright", 
       legend = c("KM (unweighted)", "Avg of Mixed Exps", "Avge of Mixed Exps2"), 
       col = c("black", "purple", "blue"))
```

Short Duration | Long Duration
---------------|--------------
`r round(shortDur/12, 2)` | `r round(longDur/12, 2)`

Age Category | Proportion Rels Short Duration
-------------|--------------
15-19 | `r round(wmix15, 2)`
20-24 | `r round(wmix20, 2)`
25-29 | `r round(wmix25, 2)`
30-34 | `r round(wmix30, 2)`
35-40 | `r round(wmix35, 2)`
40-44 | `r round(wmix40, 2)`






