# Duration Only Models {#dur}

This is essentially what I completed for CSSS 544 - event history analysis final project

Standard method for dissolution model in our ERGMs: homogenous constant hazard (exponential) OR, within-group homogenous constant hazard

Motivating question: Is there evidence of a non-constant hazard in empirical data? 

Method: Compare various covariate-free models with different distributions visually to Modified Kaplan-Meier. 

## Plots 
```{r load-fits, echo = FALSE}
expFit <- readRDS(here("model_fits", "nocovs", "exp.rds")) 
#km <- readRDS(here("model_fits", "km", "km-nsfg.rds"))
#km_weighted <- readRDS(here("model_fits", "km", "km-weighted.rds"))
weibFit <- readRDS(here("model_fits", "nocovs", "weib.rds"))
gammaFit <- readRDS(here("model_fits", "nocovs", "gamma.rds"))
mixExpFit <- readRDS(here("model_fits", "mixed", "mix-exp-nsfg.rds"))
mixExpFitweighted <- readRDS(here("model_fits", "mixed", "mix-exp-nsfg-weighted.rds"))

# functions
logit = function(p){
  log(p/(1-p))
}

invLogit = function(phi){
  1/(1+exp(-phi))
}
```
```{r survival-plots, echo = FALSE}
# full plot
durVec <- c(1:max(km_weighted$time))

plot(km_weighted, 
     conf.int = FALSE,
     xlab = "t, months", ylab = "S(t)",
     main="Survival of Relationships"
)
lines(expFit, col="blue", lty=2)
lines(weibFit, col="red", lwd=2)
#lines(gammaFit, col="darkgreen")
lines(x = durVec,
      y = invLogit(coef(mixExpFitweighted)[3]) *
      pexp(q = durVec, exp(coef(mixExpFitweighted)[1]), lower.tail = FALSE) +
      (1-invLogit(coef(mixExpFitweighted)[3])) *
      pexp(q = durVec, exp(coef(mixExpFitweighted)[2]), lower.tail = FALSE),
      col ="purple", lwd=2
)
legend("topright", legend = c("Kaplan-Meier", "Exponential", "Mixed Exponential", "Weibull"),
col = c("black", "blue", "purple", "red"), cex=0.7, lwd=2, lty = c(1,2,1,1))


# zoomed plot
plot(km_weighted, 
     conf.int = FALSE,
     xlab = "t, months", ylab = "S(t)",
     xlim = c(0,150),
     main="Survival of Relationships, Zoomed"
)
lines(expFit, col="blue")
lines(weibFit, col="red")
#lines(gammaFit, col="darkgreen")
lines(x = durVec,
      y = invLogit(coef(mixExpFitweighted)[3]) *
      pexp(q = durVec, exp(coef(mixExpFitweighted)[1]), lower.tail = FALSE) +
      (1-invLogit(coef(mixExpFitweighted)[3])) *
      pexp(q = durVec, exp(coef(mixExpFitweighted)[2]), lower.tail = FALSE),
      col ="purple", lwd=2
)
legend("topright", legend = c("Kaplan-Meier", "Exponential", "Mixed Exponential", "Weibull"),
col = c("black", "blue", "purple", "red"), cex=0.7, lwd=2)
```




```{r mix-exp-km, echo=FALSE, include=FALSE}
plot(km_weighted, 
     conf.int = FALSE,
     xlab = "t, months", ylab = "S(t)",
     main="Mixed Exponential, No Covs"
)
lines(x = durVec,
      y = invLogit(coef(mixExpFit)[3]) *
      pexp(q = durVec, exp(coef(mixExpFit)[1]), lower.tail = FALSE) +
      (1-invLogit(coef(mixExpFit)[3])) *
      pexp(q = durVec, exp(coef(mixExpFit)[2]), lower.tail = FALSE),
      col ="purple", lwd=2
)
legend(300,1, legend = c("Kaplan-Meier (unweighted)", "Mixed Exp"),
col = c("black", "purple"), cex=0.7, lwd=2)

```


## Model Comparison


Model | AIC
---------- | ---------------------
Exponential | `r round(AIC(expFit))`
Weibull | `r round(AIC(weibFit))`
Gamma | `r round(AIC(gammaFit))`


## Limitations / Future Work  
 - Statistical estimator - Kirk uses Kolmogorov-Smirnov Distance
 - Random Effects model for serial observations? 
    is this necessary? In Burington et al. proposed modified K-M, partner subsampling is only addressed in the way they calculate the CIs (using a boostrapping technique that I haven't employed yet) If the reference K-M considers all rels as independent, then shouldn't my exploratory models? 
