# Younger Population {#young}

This section explores models fit to the 15-29 year olds in NSFG, my population of interest to model in chapters 2 & 3 of dissertation. 

Will likely stick to mixture model framework for network model in chapters 2 and 3. 
```{r load-young-data, echo=FALSE}
##### kaplan-meier ######
# no covs 
km_unweighted <- readRDS(here("model_fits","1529", "km-unweighted.rds"))
km_weighted <- readRDS(here("model_fits", "1529", "km-weighted.rds"))
# covs
km_ac <- readRDS(here("model_fits","1529", "km_ac_weighted.rds"))
km_iac <- readRDS(here("model_fits","1529", "km_i_agecat_weighted.rds"))
km_reltype <- readRDS(here("model_fits","1529", "km_reltype_weighted.rds"))
km_partsyr <- readRDS(here("model_fits","1529", "km_eparts_weighted.rds"))
km_race <- readRDS(here("model_fits","1529", "km_erace_weighted.rds"))


####### Parametric 
# no covs for reference
exp.young <- readRDS(here("model_fits", "1529","expFlex.rds"))
weib.young <- readRDS(here("model_fits", "1529", "weibFlex.rds"))
# mix.exp.young 

# same covs as full data set 
# univariate 
###### exponential 
e.race <- readRDS(here("model_fits", "1529", "erace.rds")) 
e.reltype <- readRDS(here("model_fits", "1529", "ereltype.rds")) 
e.partsyr <- readRDS(here("model_fits", "1529", "epartsyr.rds"))
e.network <- readRDS(here("model_fits", "1529", "enetwork.rds"))
e.agecat <- readRDS(here("model_fits", "1529", "eagecat.rds"))

###### weibull 
w.race <- readRDS(here("model_fits", "1529", "wrace.rds")) 
w.reltype <- readRDS(here("model_fits", "1529", "wreltype.rds")) 
w.partsyr <- readRDS(here("model_fits", "1529", "wpartsyr.rds"))
w.network <- readRDS(here("model_fits", "1529", "wnetwork.rds"))
w.agecat <- readRDS(here("model_fits", "1529", "wagecat.rds"))


# multivariate 
###### exponential 
e.raceage <- readRDS(here("model_fits", "1529", "eraceage.rds")) 

e.racerels <- readRDS(here("model_fits", "1529", "eracerels.rds"))
e.agerels <- readRDS(here("model_fits", "1529", "eagerels.rds"))

e.ageparts <- readRDS(here("model_fits", "1529", "eageparts.rds")) 
e.raceparts <- readRDS(here("model_fits", "1529", "eraceparts.rds")) 

e.agedeg <- readRDS(here("model_fits", "1529", "eagedeg.rds")) 
e.racedeg <- readRDS(here("model_fits", "1529", "eracedeg.rds"))

e.relparts <- readRDS(here("model_fits", "1529", "erelparts.rds"))

###### weibull 
w.raceage <- readRDS(here("model_fits", "1529", "wraceage.rds"))

w.racerels <- readRDS(here("model_fits", "1529", "wracerels.rds"))
w.agerels <- readRDS(here("model_fits", "1529", "wagerels.rds"))

w.ageparts <- readRDS(here("model_fits", "1529", "wageparts.rds")) 
w.raceparts <- readRDS(here("model_fits", "1529", "wraceparts.rds")) 

#w.agedeg <- readRDS(here("model_fits", "1529", "wagedeg.rds")) 
w.racedeg <- readRDS(here("model_fits", "1529", "wracedeg.rds"))

w.relparts <- readRDS(here("model_fits", "1529", "wrelparts.rds"))
```

## Kaplan-Meier Descriptives
```{r km-young, echo=FALSE}
durVec <- c(1:max(km_weighted$time))

fits <- list(KMweighted = km_weighted, KM = km_unweighted)
ggsurvplot(fits, data=dat, combine=TRUE, ggtheme=theme_gray())
```


### Current AgeCat
```{r km-agecat-current-young, echo=FALSE}
autoplot(km_ac, censor=FALSE, conf.int=FALSE,
     main="Weighted K-M Survival Curves by Current AgeCat",
     ylab="S(t)", xlab="t, months")
```

### Initial AgeCat
```{r km-agecat-initial-young, echo=FALSE}
autoplot(km_iac, censor=F,
     main="Weighted K-M Survival Curves by Initial AgeCat",
     ylab="S(t)", xlab="months")
```

### Relationship Type
```{r reltype-young, echo=FALSE}
autoplot(km_reltype, censor=F, conf.int=FALSE,
     main="Weighted K-M Survival Curves by Relationship Type",
     ylab="S(t)", xlab="months")
```

### Number Partners in Last Year 
topcoded at 3; vaginal sex partners 

```{r partsyr3-young, echo=FALSE}
autoplot(km_partsyr, censor=F, conf.int=FALSE,
     main="Weighted K-M Survival Curves by Ego's Partners in Last Year",
     ylab="S(t)", xlab="months")
```

### Race 
```{r km-erace-young, echo=FALSE}
autoplot(km_race, censor=F, conf.int=FALSE,
     main="Weighted K-M Survival Curves by Ego's Race",
     ylab="S(t)", xlab="months")
```

## Parametric Models

### Age Category (Current)
```{r agecat-young, echo=FALSE, warning=FALSE}
s <- summary(e.agecat)
plot(km_ac, lty=2, col = c(2:6,"orange"), lwd=2,
     ylab="S(t)", xlab="Months", 
     main="Current Age Category, Exponential Dist")
lines(s$`e.agecat=15-19`$est, type = "lines", col = 2, lwd=1.5)
lines(s$`e.agecat=20-24`$est, type = "lines", col = 3, lwd=1.5)
lines(s$`e.agecat=25-29`$est, type = "lines", col = 4, lwd=1.5)
legend("topright", lty=c(2,1,1,1,1), lwd=c(rep(2,5)), col=c(1,1,2,3,4), 
       legend = c("K-M", "Fitted Model", "15-19", "20-24", "25-29"), cex = 0.8, bty="n")

s <- summary(w.agecat)
plot(km_ac, lty=2, col = c(2:6,"orange"), lwd=2,
     ylab="S(t)", xlab="Months", 
     main="Current Age Category, Weibull Dist")
lines(s$`e.agecat=15-19`$est, type = "lines", col = 2, lwd=1.5)
lines(s$`e.agecat=20-24`$est, type = "lines", col = 3, lwd=1.5)
lines(s$`e.agecat=25-29`$est, type = "lines", col = 4, lwd=1.5)
legend("topright", lty=c(2,rep(1,4)), lwd=c(rep(2,8)), col=c(1,1,2,3,4), 
       legend = c("K-M", "Fitted Model", "15-19", "20-24", "25-29"), cex = 0.8, bty="n")
```

### Race

```{r race-cov-young, echo=FALSE}
plot(e.race, main = "Ego Race, Exp Dist", 
     col=c("red", "blue", "springgreen3", "orange"), xlab="t, months")
legend(370, 0.95, legend = c("black", "white", "hisp", "other"), col=c("blue", "red", "springgreen3", "orange"), lwd=2, lty=1, cex=.7)
plot(w.race, main = "Ego Race, Weibull Dist", 
     col=c("red", "blue", "springgreen3", "orange"), xlab="t, months")
legend(370, 0.95, legend = c("black", "white", "hisp", "other"), col=c("blue", "red", "springgreen3", "orange"), lwd=2, lty=1, cex=.7)
```

### Rel Type
```{r reltype-cov-young, echo=FALSE}
plot(e.reltype, main = "Rel Type, Exp Dist", 
     col=c("red", "blue", "springgreen3"), xlab="t, months")
legend(370, 1, legend = c("Marriage", "Cohab", "Other"), col=c("red", "springgreen3", "blue"), lwd=2, lty=1, cex=.7)

plot(w.reltype, main = "Rel Type, Weibull Dist", 
     col=c("red", "blue", "springgreen3"), xlab="t, months")
legend(370, 1, legend = c("Marriage", "Cohab", "Other"), col=c("red", "springgreen3", "blue"), lwd=2, lty=1, cex=.7)
```

### Network  
Similar to above, but split into marcoh and other - two of the networks we often split out
```{r network-cov-young, echo=FALSE}
plot(e.network, main = "Network, Exp Dist", 
     col=c("red", "blue"), xlab="t, months")
legend(370, 1, legend = c("Mar/Coh", "Other"), col=c("red","blue"), lwd=2, lty=1, cex=.7)

plot(w.network, main = "Network, Weibull Dist", 
     col=c("red", "blue"), xlab="t, months")
legend(370, 1, legend = c("Mar/Coh", "Other"), col=c("red","blue"), lwd=2, lty=1, cex=.7)
```

### Partners in Last Year

Note: this was also run with "osnpyr" and "maxospyr" with very little differences (not shown)  

```{r partsyr-cov-young, echo=FALSE}
plot(e.partsyr, main = "Partners in Last Year, Exp Dist", 
     col=c("red", "blue", "springgreen3"), xlab="t, months")
legend(370, 1, legend = c("1", "2", "3+"), col=c("red", "springgreen3", "blue"), lwd=2, lty=1, cex=.7)

plot(w.partsyr, main = "Partners in Last Year, Weibull Dist", 
     col=c("red", "blue", "springgreen3"), xlab="t, months")
legend(370, 1, legend = c("1", "2", "3+"), col=c("red", "springgreen3", "blue"), lwd=2, lty=1, cex=.7)
```

### Momentary Degree
```{r deg-cov-young, echo=FALSE}
plot(e.deg, main = "Momentary Degree, Exp Dist",
     col=2:5, xlab="t, months")
legend("right", 1, legend = c("0", "1", "2", "3"), col=c(3,2,5,4), lwd=2, lty=1, cex=.7)

plot(w.deg, main = "Momentary Degree, Weibull Dist",
     col=2:5, xlab="t, months")
legend("right", 1, legend = c("0", "1", "2", "3"), col=c(3,2,5,4), lwd=2, lty=1, cex=.7)
```


### AIC Comparison 

Model | Exp Dist | Weibull 
------|----------|--------
No Covs | `r AIC(expFit)` | `r AIC(weibFit)`
Age Category |  `r AIC(e.agecat)` |  `r AIC(w.agecat)`
Race | `r AIC(e.race)` | `r AIC(w.race)`
Rel Type - 3 |  `r AIC(e.reltype)` | `r AIC(w.reltype)`
Rel Type - 2 | `r AIC(e.network)` | `r AIC(w.network)`
Parts in Last Year | `r AIC(e.partsyr)` | `r AIC(w.partsyr)`
Momentary Degree | `r AIC(e.deg)` | `r AIC(w.deg)` 




## Latent Mixture Models

__NOTE: these mixture models do NOT use NSFG survey weights__

```{r young-mix-models, echo=FALSE, eval=FALSE}

durVec <- 1:max(km_weighted$time)

######################################################
# mixture model, no covariates
######################################################
mixExpYoung <- readRDS(here("model_fits", "1529", "mixExpYoung.rds"))

l1 <- exp(coef(mixExpYoung)[1])
l2 <- exp(coef(mixExpYoung)[2])
weight <- invLogit(coef(mixExpYoung)[3])

mixYoung <- weight * pexp(q = durVec, l1, lower.tail = FALSE) +
            (1-weight) * pexp(q = durVec, l2, lower.tail = FALSE)
        

######################################################
# mixture model, lambda weights are variable 
######################################################
weightsMixExpyoung <- readRDS(here("model_fits", "1529", "weightsMixExp-young.rds"))

shortDur <- 1/exp(coef(weightsMixExpyoung)[1])
longDur <- 1/exp(coef(weightsMixExpyoung)[2])

wmix15 <- invLogit(coef(weightsMixExpyoung)[3])
wmix20 <- invLogit(coef(weightsMixExpyoung)[4])
wmix25 <- invLogit(coef(weightsMixExpyoung)[5])


dfwYoung <- as.data.frame(cbind(
            durVec, 
            wmix15 * pexp(q = durVec, exp(coef(weightsMixExpyoung)[1]), lower.tail = FALSE) +
        (1-wmix15) * pexp(q = durVec, exp(coef(weightsMixExpyoung)[2]), lower.tail = FALSE),
            wmix20 * pexp(q = durVec, exp(coef(weightsMixExpyoung)[1]), lower.tail = FALSE) +
        (1-wmix20) * pexp(q = durVec, exp(coef(weightsMixExpyoung)[2]), lower.tail = FALSE), 
            wmix25 * pexp(q = durVec, exp(coef(weightsMixExpyoung)[1]), lower.tail = FALSE) +
        (1-wmix25) * pexp(q = durVec, exp(coef(weightsMixExpyoung)[2]), lower.tail = FALSE)))

colnames(dfwYoung) <- c("t", "15-19", "20-24", "25-29")

# weighted average model results 
p <- dat %>% count(e.agecat) %>% mutate(prop = n/sum(n))

dfwYoung$w15 <- as.numeric(rep(p[1,3], length(dfwYoung$t)))
dfwYoung$w20 <- as.numeric(rep(p[2,3], length(dfwYoung$t)))
dfwYoung$w25 <- as.numeric(rep(p[3,3], length(dfwYoung$t)))

mix1young <- as.data.frame(cbind(durVec, dfwYoung$`15-19`, dfwYoung$`20-24`, dfwYoung$`25-29`))

avg_wYoung1 <- dfwYoung$`15-19`*dfwYoung$w15 + 
              dfwYoung$`20-24`*dfwYoung$w20 + 
              dfwYoung$`25-29`*dfwYoung$w25

######################################################
# mixture model, lambdas and lambda weights are variable 
######################################################
dfwYoung2 <- (df[1:192,1:4])
avg_wYoung2 <- dfwYoung2$`15-19`*dfwYoung$w15 + 
               dfwYoung2$`20-24`*dfwYoung$w20 + 
               dfwYoung2$`25-29`*dfwYoung$w25

#######################################################
# plots
#######################################################

# make data frame 
dfYoung <- as.data.frame(cbind(km_unweighted$surv, avg_wYoung1, avg_wYoung2, mixYoung, durVec))
colnames(dfYoung) <- c("KM", "Mix1", "Mix2" , "NoCovs", "t")

plot_ly(dfYoung, x=~t, y=~KM, type="scatter", mode="lines", name="K-M") %>%
    add_trace(y=~NoCovs, mode="lines", name = "Mixture, No Covs") %>%
    add_trace(y=~Mix2, mode="lines", name = "Mixture by AgeCat") %>%
    layout(title="K-M versus Mixture Models (no survey weights)",
         yaxis = list(title="S(t)"), 
         xaxis= list("t, months"))
```

## Other 


```{r exp-young-plots-other, echo=FALSE}
####### load other age models ##############
# current ego age 
expYoungCurrAge <- readRDS(here("model_fits", "1529", "old", "expYoungCurrAge.rds"))
weibYoungCurrAge <- readRDS(here("model_fits", "1529", "old", "weibYoungCurrAge.rds"))

# sqrt age diff 
expYoungAgeDiff <- readRDS(here("model_fits", "1529", "old", "expYoungAgeDiff.rds"))
weibYoungAgeDiff <- readRDS(here("model_fits", "1529", "old", "weibYoungAgeDiff.rds"))

# age cat 
expYoungAgeCat <- readRDS(here("model_fits", "1529", "old", "expYoungAgeCat.rds"))
weibYoungAgeCat <- readRDS(here("model_fits", "1529", "old", "weibYoungAgeCat.rds"))

maxLength <- 
##### plot 
plot(exp.young,
     main="Young Pop, Exponential",
     ylab="S(t)", xlab="t, months", lwd=3,
     xlim=c(0, 250))
lines(expYoungAgeDiff, col="springgreen3", lwd=3)
lines(expYoungCurrAge,col="steelblue3", lwd=3)
legend(150,1, legend=c("K-M", "No Covs", "Age Diff", "Current Ego Age"),
       col=c(1, "red", "springgreen3", "steelblue3"), lwd=3)

print("all models have same curve")

plot(weib.young,
     main="Young Pop, Weibull",
     ylab="S(t)", xlab="t, months", lwd=3,
      xlim=c(0, 250))
lines(weibYoungAgeDiff, col="springgreen3", lwd=3)
lines(weibYoungCurrAge,col="steelblue3", lwd=3)
legend(150,1, legend=c("K-M", "No Covs", "Age Diff", "Current Ego Age"),
       col=c(1, "red", "springgreen3", "steelblue3"), lwd=3)
```