---
title: "More Data for Duration Sims"
author: "Emily Pollock"
date: "10/28/2019"
output: 
  html_document:
    toc: true
    toc_float: true
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(plotly)
library(srvyr)
library(reshape2)
library(kableExtra)
library(ergm.ego)
#library(scales)
#library(DataCombine)
library(DT)
```
```{r load-data, echo=FALSE}
nsfg <- readRDS("~/NSFG_DATA/Full/nsfg_complete.rds") 
nsfg <- nsfg %>%
  mutate(sex = as.factor(sex), 
         age = as.factor(age), 
         hadsex = as.factor(HADSEX),
         ssany = as.factor(ifelse(SAMESEXANY==1, 1, 2)),
         sexage = ifelse(agefsex <= 14, 14, 
                  ifelse(agefsex > 14, agefsex, NA)),
         sexage = replace_na(sexage, 100))


svy <- as_survey_design(nsfg, weights = weight, ids = ego)

svy_unw <- as_survey_design(nsfg, weights = 1, ids = ego)

#  egodata
egodat <- readRDS("~/NSFG_DATA/Objects/fullEgodata.rds")
```

## Sexual Debut
***  
### Opposite-Sex Contact {.tabset .tabset-pills .tabset-fade}
(var = hadsex)

#### Overall {.tabset .tabset-pills .tabset-fade}
##### Ever
```{r debut, echo=FALSE}
debut <- svy %>% 
          group_by(sex, age, hadsex) %>% 
          summarize(prop_debut = survey_mean()) %>%
          mutate(prop_debut=round(prop_debut,3)) %>%
          filter(hadsex==1) %>%
          select(-hadsex, -prop_debut_se)

p <-ggplot(debut, aes(x=age, y=prop_debut, group=sex)) +
          geom_line(aes(color=sex)) +
          labs(title="Proportion Debuted w/ Opposite Sex")
  
ggplotly(p) 

# fit poisson curve to the data 

# females
#f1 <- glm(prop_debut ~ age, family = "logistic", data=debut[debut$sex=="F",])

#summary(m1 <- glm(num_awards ~ prog + math, family="logistic", data=p))
```

##### In Last Year 
```{r last-year, echo=FALSE}
lastyr <- svy %>% 
          mutate(osyr = as.factor(ifelse(osnpyr3 == 0, 0, 1))) %>%
          group_by(sex, age, osyr) %>% 
          summarize(prop_osyr = survey_mean()) %>%
          mutate(prop_osyr=round(prop_osyr,3)) %>%
          filter(osyr==1) %>%
          select(-osyr, -prop_osyr_se)

p <-ggplot(lastyr, aes(x=age, y=prop_osyr, group=sex)) +
          geom_line(aes(color=sex)) +
          labs(title="Proportion Had Sex w/ Opposite Sex in Last Year, ospyn")
  

ggplotly(p) 

lastyrprts <- svy %>% 
          mutate(partsyr3 = ifelse(is.na(partsyr3), 0, partsyr3)) %>%
          mutate(osyr = as.factor(ifelse(partsyr3 == 0, 0, 1))) %>%
          group_by(sex, age, osyr) %>% 
          summarize(prop_osyr = survey_mean()) %>%
          mutate(prop_osyr=round(prop_osyr,3)) %>%
          filter(osyr==1) %>%
          select(-osyr, -prop_osyr_se)

p <-ggplot(lastyrprts, aes(x=age, y=prop_osyr, group=sex)) +
          geom_line(aes(color=sex)) +
          labs(title="Proportion Had Sex w/ Opposite Sex in Last Year, partsyr3")
  

ggplotly(p) 
```

##### Comparison
```{r compare, echo=FALSE}
both <- cbind(debut, lastyr[,3])
bothF <- both %>%
          filter(sex == "F") %>%
          select(-sex) %>%
          gather('prop_debut', 'prop_osyr', key = "type", value="prop")

bothM <- both %>%
          filter(sex == "M") %>%
          select(-sex) %>%
          gather('prop_debut', 'prop_osyr', key = "type", value="prop")


p1 <- ggplot(bothF, (aes(x=age, y=prop, group=type))) + 
        geom_line(aes(color=type)) +
        ggtitle("Females, with ospyn vars")

p2 <- ggplot(bothM, (aes(x=age, y=prop, group=type))) + 
        geom_line(aes(color=type)) +
        ggtitle("Males, with ospyn")

ggplotly(p1);ggplotly(p2)

both2 <- cbind(debut, lastyrprts[,3])
bothF2 <- both2 %>%
          filter(sex == "F") %>%
          select(-sex) %>%
          gather('prop_debut', 'prop_osyr', key = "type", value="prop")

bothM2 <- both %>%
          filter(sex == "M") %>%
          select(-sex) %>%
          gather('prop_debut', 'prop_osyr', key = "type", value="prop")


p12 <- ggplot(bothF2, (aes(x=age, y=prop, group=type))) + 
        geom_line(aes(color=type)) +
        ggtitle("Females, with partsyr3")

p22 <- ggplot(bothM2, (aes(x=age, y=prop, group=type))) + 
        geom_line(aes(color=type)) +
        ggtitle("Males, with partsyr3")

ggplotly(p12);ggplotly(p22)

```

#### By Birth Cohort
```{r debut2, echo=FALSE}

debut2 <- svy %>% 
          group_by(sex, birthcohort, hadsex) %>% 
          summarize(proportion = survey_mean()) %>%
          mutate(proportion=round(proportion,3)) %>%
          filter(hadsex==1)

pd <-ggplot(debut2, aes(x=birthcohort, y=proportion, group=sex)) +
          geom_line(aes(color=sex)) +
          labs(title="Proportion Debuted w/ Opposite Sex")
  

ggplotly(pd) 

```

### Same Sex Contact {.tabset .tabset-pills .tabset-fade}

#### Overall

```{r sscontact-1, echo=FALSE, fig.align="center"}
# overall average
ssavg <- svy %>% 
          filter(!is.na(ssany)) %>%
          mutate(sexhist = as.factor(paste(hadsex,ssany, sep = ""))) %>% 
          mutate(sexhist = factor(sexhist, labels = c("OS+SS", "OS Only", "SS Only", "No Sex"))) %>%
          group_by(sex, sexhist) %>% 
          summarize(proportion = survey_mean()) %>%
          mutate(proportion=round(proportion,3),
                 proportion_se=round(proportion_se,3)) %>%
          filter(sexhist %in% "SS Only" | sexhist %in% "OS+SS") 

datatable(ssavg)
```

#### By Sex
```{r ss-age, echo=FALSE}
# by age
ssany <- svy %>% 
          filter(!is.na(ssany)) %>%
          group_by(sex, age, ssany) %>% 
          summarize(proportion = survey_mean()) %>%
          mutate(proportion=round(proportion,3)) %>%
          filter(ssany==1)

p2 <- ggplot(ssany, aes(x=age, y=proportion, fill=sex)) +
        geom_col() +
        labs(title="Proportion Reporting Any Same Sex Contact") + 
        scale_x_discrete(breaks=seq(0,45,5)) +
        facet_wrap(vars(sex))

ggplotly(p2)
```

#### By Birth Cohort

```{r sscontact-2, echo=FALSE, fig.align="center"}
ssbc <- svy %>% 
          filter(!is.na(ssany)) %>%
          group_by(sex, birthcohort, ssany) %>% 
          summarize(proportion = survey_mean()) %>%
          mutate(proportion=round(proportion,3)) %>%
          filter(ssany==1)

p3 <- ggplot(ssbc, aes(x=birthcohort, y=proportion, fill=sex)) +
        geom_col() +
        labs(title="Proportion Reporting Any Same Sex Contact", subtitle = "By Birth Cohort") + 
        facet_wrap(vars(sex))

ggplotly(p3)
```

#### By Opposite-Sex Contact

```{r ss-3, echo=FALSE}
ssos <- svy %>% 
          filter(!is.na(ssany)) %>%
          mutate(sexhist = as.factor(paste(hadsex,ssany, sep = ""))) %>% 
          mutate(sexhist = factor(sexhist, labels = c("OS+SS", "OS Only", "SS Only", "No Sex"))) %>%
          group_by(sex, age, sexhist) %>% 
          summarize(proportion = survey_mean()) %>%
          mutate(proportion=round(proportion,3)) 


#
            
            
p4 <- ggplot(ssos, aes(x=age, y=proportion, fill=sexhist)) +
        geom_col() +
        scale_x_discrete(breaks=seq(0,45,5)) +
        labs(title="Sex Experience", subtitle = "By Age") + 
        facet_wrap(vars(sex))

ggplotly(p4)

# labels = c("OS+SS", "OS Only", "SS Only", "No Sex")

ssos2 <- svy %>% 
          filter(!is.na(ssany)) %>%
          mutate(sexhist = as.factor(paste(hadsex,ssany, sep = ""))) %>% 
          mutate(sexhist = factor(sexhist, labels = c("OS+SS", "OS Only", "SS Only", "No Sex"))) %>%
          group_by(sex, birthcohort, sexhist) %>% 
          summarize(proportion = survey_mean()) %>%
          mutate(proportion=round(proportion,3)) 

p5 <- ggplot(ssos2, aes(x=birthcohort, y=proportion, fill=sexhist)) +
        geom_col() +
        labs(title="Sex Experience", subtitle = "By Birth Cohort") + 
        facet_wrap(vars(sex))

ggplotly(p5)
```


#### SS Contact Only 

```{r ss-4, echo=FALSE}
# by age
sso <- svy %>% 
          filter(!is.na(ssany)) %>%
          mutate(sexhist = as.factor(paste(hadsex,ssany, sep = ""))) %>% 
          mutate(sexhist = factor(sexhist, labels = c("OS+SS", "OS Only", "SS Only", "No Sex"))) %>%
          group_by(sex, age, sexhist) %>% 
          summarize(proportion = survey_mean()) %>%
          mutate(proportion=round(proportion,3)) %>%
          filter(sexhist %in% "SS Only")
            
p6 <- ggplot(sso, aes(x=age, y=proportion, fill=sex)) +
        geom_col() +
        scale_x_discrete(breaks=seq(0,45,5)) +
        labs(title="SS-Only Experience", subtitle = "By Age") + 
        facet_wrap(vars(sex))

ggplotly(p6)

#by birth cohort 
ssobc <- svy %>% 
          filter(!is.na(ssany)) %>%
          mutate(sexhist = as.factor(paste(hadsex,ssany, sep = ""))) %>% 
          mutate(sexhist = factor(sexhist, labels = c("OS+SS", "OS Only", "SS Only", "No Sex"))) %>%
          group_by(sex, birthcohort, sexhist) %>% 
          summarize(proportion = survey_mean()) %>%
          mutate(proportion=round(proportion,3)) %>%
          filter(sexhist %in% "SS Only")
            
p7 <- ggplot(ssobc, aes(x=birthcohort, y=proportion, fill=sex)) +
        geom_col() +
        scale_x_discrete(breaks=seq(0,45,5)) +
        labs(title="SS-Only Experience", subtitle = "By Birth Cohort") + 
        facet_wrap(vars(sex))

ggplotly(p7)
```





```{r het-last-year, echo=FALSE}



```


## Sexual Debut - Probability by Age 

age 14 = those who debuted before or at age 14
age 100 = those who have never debuted w/ opposite sex

go back and remind myself how "agefsex"

```{r debut-by-age, echo=FALSE}
# what we want is not the cumulative probabiltiy that someone in the population has debuted by a certain age, but instead the probability that a person debuts AT a certain age 

debut_age <- svy %>% 
          mutate(sexage = as.factor(sexage)) %>%
          group_by(sex, sexage) %>% 
          summarize(proportion = survey_mean()) %>%
          mutate(proportion=round(proportion,3)) 

pda <-ggplot(debut_age, aes(x=sexage, y=proportion, group=sex)) +
          geom_line(aes(color=sex)) +
          labs(title="Proportion Who Debuted w/ Opposite Sex at Certain Age")

ggplotly(pda) 

debut_age2 <- debut_age %>% 
  select(-proportion_se) %>%
  spread(sex, proportion)

datatable(debut_age2)
```
## Probs for model 
```{r convert-to-weeks, echo=FALSE}
# probability that someone debutes that year by age & sex
prob_debut_year <- debut_age2[,2:3]

adjDebutTable <- matrix(c(prob_debut_year$F[1:16], rep(0.001, 14),
                          prob_debut_year$M[1:16], rep(0.001, 14)), byrow=T, nrow=2)

# BUT now I need a different denominator - aka the number of people who hadn't debuted going into that age
# i.e. 10% of all 16 year olds debut at age 16 but what is the proportion of 16 year olds that debut GIVEN that they were 15 year olds who didn't

full_denoms <- svy %>% group_by(sex, age) %>% summarize(count = survey_total()) %>% select(count)

# 10%*all16 = x*15nosex
# x = (10%*all16)/15nosex

debuted_denoms <- svy %>% filter(hadsex==1) %>% group_by(sex, age) %>% 
  summarize(count = survey_total()) %>% select(count)

allF <- full_denoms$count[1:30]
allM <- full_denoms$count[31:60]

debutedF <- debuted_denoms$count[1:30]
debutedM <- debuted_denoms$count[31:60]

nosexF <- allF-debutedF
nosexM <- allM-debutedM
 
adjprob <- function(debutvec, fullDenoms, nosexDenoms){
  x <- NULL
  for (i in 2:length(fullDenoms)){
    p <- debutvec[i]*fullDenoms[i]/nosexDenoms[i-1]
    x <- c(x, p)  
  }
  return(x)
}

prob1644_F <- adjprob(adjDebutTable[1,], allF, nosexF)
prob1644_M <- adjprob(adjDebutTable[2,], allM, nosexM)
prob1644 <- matrix(c(prob1644_F, prob1644_M), byrow=T, nrow=2)

allprobs[,1] <- adjDebutTable[,1]
allprobs[,2:30] <- prob1644

# probability that someone debuts per week
prob_debut_week <- 1-(1-allprobs)^(1/52)
```




## Age Dist of Never Sex  
```{r neversexdist, echo=FALSE, eval=FALSE}
neverdebut_age <- svy %>% 
          filter(HADSEX == 0) %>% 
          group_by(sex, age, na.rm=F) %>% 
          summarize(proportion = survey_mean()) %>%
          mutate(proportion=round(proportion,3)) 

npda <-ggplot(neverdebut_age, aes(x=age, y=proportion, group=sex)) +
          geom_line(aes(color=sex)) +
          labs(title="Proportion Who Have Never Debuted w/ Opposite Sex at Certain Age")

ggplotly(npda) 
```

## Distribution at Cross-Section  
***  
what proportion of all rels are marriages, cohabs, others at cross-section?

```{r prop-rels, echo=FALSE}
actives <- egodat$altersAllActive

aSvy <- as_survey_design(actives, weights = weight, ids=ego)

rels <- aSvy %>% 
  group_by(reltype) %>% 
  summarize(prop = survey_mean()) %>% 
  mutate(prop = round(prop,3),
         prop_se = round(prop_se, 5))

datatable(rels, caption = "Proportion of RelType - Active Rels")
```


## Relationship Transitions {.tabset .tabset-pills .tabset-fade}
***  
### Other to Cohab {.tabset .tabset-pills .tabset-fade}

time between rel start and cohabitation

#### Current Cohabs or Marriages 

(marraiges who cohabitated premaritally)  
```{r rt-currentcohabsandmarrs, echo=FALSE, message=FALSE, warning=FALSE}
egosCM <- egodat$egos[which(egodat$egos$ego %in% egodat$altersMarCoh$ego),]

marcohs <- merge(egodat$altersMarCoh, egosCM[,c("ego", "CMINTVW", "cmcohab")], by="ego")

marcohs <- marcohs %>% 
            filter(!is.na(cmcohab)) %>%
            mutate(cmstart = CMINTVW-dfs) %>%
            mutate(timetoCohab = cmcohab-cmstart) %>%
            mutate(preCohabSex = ifelse(timetoCohab>0, "yes", 
                       ifelse(timetoCohab==0, "at cohab",
                       ifelse(timetoCohab <0, "no", NA))))


marcohSvy <- as_survey_design(marcohs, weights = weight, ids=ego)

#histogram of raw #s
#marcohs %>%
#  ggplot(aes(x = timetoCohab/12, weights = weight)) +
#  geom_histogram(binwidth = 1/12) +
#  scale_x_continuous(name="years", limits = c(-10,10)) +
#  ggtitle("Time to Cohabitation from Initiation of Sexual Contact", subtitle = "bin width = 1 month") 

marcohs %>%
  ggplot(aes(x = timetoCohab/12, weights = weight)) +
  geom_density() +
  scale_x_continuous(name="years", limits = c(-10,10)) +
  ggtitle("Time to Cohabitation from Initiation of Sexual Contact", subtitle = "bin width = 1 month") 

# other to cohab (either current cohab or current marriage)
otc <- marcohSvy %>% summarize(meanYears = survey_mean(timetoCohab/12, na.rm=T))

# if you limit to only those who report positive
otcp <- marcohSvy %>% filter(timetoCohab > 0) %>%
  summarize(meanYears = survey_mean(timetoCohab/12, na.rm=T))

# if you limit to only those who report negative
otcn <- marcohSvy %>% filter(timetoCohab < 0) %>%
  summarize(meanYears = survey_mean(timetoCohab/12, na.rm=T))

otc_dt <- rbind(otc, otcp, otcn) 
rownames(otc_dt) <- c("All Other->Cohab", "Only Positive Time", "Only Negative Time")
otc_dt <- round(otc_dt, 3)

datatable(otc_dt, caption = "Mean time to cohabitation")

# what prop report cohabitating prior to sex 
preCsex <- marcohSvy %>% 
  filter(!is.na(preCohabSex)) %>%
  group_by(preCohabSex) %>%
  summarize(prop=survey_mean()) %>%
  mutate(prop = round(prop,3),
         prop_se = round(prop,3))

datatable(preCsex, caption = "Do you report sex before cohab?")
```


#### Current Cohabs ONLY 
```{r rt-currentcohabs, echo=FALSE, warning=FALSE}
egosCohabs <- egodat$egos[which(egodat$egos$ego %in% egodat$altersCohabs$ego),]

cohabs <- merge(egodat$altersCohabs, egosCohabs[,c("ego", "CMINTVW", "cmcohab")], by="ego")
cohabs <- cohabs %>% 
            filter(!is.na(cmcohab)) %>%
            mutate(cmstart = CMINTVW-dfs) %>%
            mutate(timetoCohab = cmcohab-cmstart) %>%
            mutate(preCohabSex = ifelse(timetoCohab>0, "yes", 
                       ifelse(timetoCohab==0, "at cohab",
                       ifelse(timetoCohab <0, "no", NA))))

cohabSvy <- as_survey_design(cohabs, weights = weight, ids=ego)

# other to cohab
otco <- cohabSvy %>% summarize(meanTime = survey_mean(timetoCohab/12, na.rm=T))

# if you limit to only those who report positive
otcop <- cohabSvy %>% filter(timetoCohab > 0) %>%
  summarize(meanTime = survey_mean(timetoCohab/12, na.rm=T))

# if you limit to only those who report negative
otcon <- cohabSvy %>% filter(timetoCohab < 0) %>%
  summarize(meanTime = survey_mean(timetoCohab/12, na.rm=T))



#histogram of raw #s
#cohabs %>%
#  ggplot(aes(x = timetoCohab/12, weights = weight)) +
#  geom_histogram(binwidth = 0.5) +
#  scale_x_continuous(name="years", limits = c(-10,10)) +
#  ggtitle("Time to Cohabitation from Initiation of Sexual Contact")

cohabs %>%
  ggplot(aes(x = timetoCohab/12, weights = weight)) +
  geom_density() +
  scale_x_continuous(name="years", limits = c(-10,10)) +
  ggtitle("Time to Cohabitation from Initiation of Sexual Contact")


otco_dt <- rbind(otco, otcop, otcon) 
rownames(otco_dt) <- c("All Other->Cohab", "Only Positive Time", "Only Negative Time")
otc_dt <- round(otc_dt, 3)

datatable(otco_dt, caption = "Mean time to cohabitation")

preCsex <- cohabSvy %>% 
  filter(!is.na(preCohabSex)) %>%
  group_by(preCohabSex) %>%
  summarize(prop=survey_mean()) %>%
  mutate(prop = round(prop,3),
         prop_se = round(prop,3))

datatable(preCsex, caption = "Do you report sex before cohab?")

```

### Cohab to Marriage 
```{r rt-cohabtomarraige, echo=FALSE}
egosmarr <- egodat$egos[which(egodat$egos$ego %in% egodat$altersMarriages$ego),]

marrs <- merge(egodat$altersMarriages, egosmarr[,c("ego", "CMINTVW", "cmcohab", "cmmarr")], by="ego")

marrs <- marrs %>% 
  mutate(pmcohab = ifelse(is.na(cmcohab), "no", "yes")) %>%
  mutate(cmstart = CMINTVW-dfs) %>%
  mutate(cohabtoMarr = cmmarr-cmcohab) %>%
  mutate(othertoMarr = cmmarr-cmstart)

marrSvy <- as_survey_design(marrs, weights = weight, ids=ego)

# how many people report cohabitating pre-maritally?
pmc <- marrSvy %>% group_by(pmcohab) %>% summarize(prop = survey_mean())

# cohab to marriage
ctm <- marrSvy %>% summarize(meanYears = survey_mean(cohabtoMarr/12, na.rm=T))

datatable(pmc, caption = "Did you cohabitate pre-maritally?")

datatable(ctm, caption = "Mean years to marriage from cohabitation")
```

```{r marrhist, echo=FALSE, warning=FALSE}
#histogram of raw #s
marrs %>%
  filter(!is.na(cohabtoMarr)) %>%
  ggplot(aes(x = cohabtoMarr/12, weights = weight)) +
  geom_density() +
  scale_x_continuous(name="years", limits = c(0,10)) +
  ggtitle("Time to Marriage from Initial Cohab")
```



### Other to Marriage
```{r rt-othertomarraige, echo=FALSE, warning=FALSE}
# other to marriage
otm <- marrSvy %>% summarize(meanYears = survey_mean(othertoMarr/12, na.rm=T))

datatable(otm, caption = "Mean year to Marriage from Non-cohabitating relationship")


#histogram of raw #s
marrs %>%
  ggplot(aes(x = othertoMarr/12, weights = weight)) +
  geom_density() +
  scale_x_continuous(name="years", limits = c(0,10)) +
  ggtitle("Time to Marriage from Initiation of Sexual Contact")
```


## Relationship Length {.tabset .tabset-pills .tabset-fade}
***

mean of active relationships (where we assume exponential dist)
plus
estimates from survival analysis (using all reported relationships), exponential model fit 

### Counterfactual Networks 

```{r reldur-cf, echo=FALSE}

expFit_cf <- readRDS(here("Survival_Analysis", "model_fits", "covs", "enetwork.rds")) 
marcoh_mean <- 1/exp(coef(expFit_cf)[[1]])
other_mean <- marcoh_mean*(1/exp(coef(expFit_cf)[[2]]))

means <- as.data.frame(rbind(marcoh_mean, other_mean))
means <- round(means/12, 2)
colnames(means) <- c("Model Predictions")
rownames(means) <- c("Marcoh", "Other")

dat <- readRDS("~/NSFG_DATA/Objects/alters_active.rds")
datsvy <- as_survey_design(dat, weights = weight, ids=ego)

means2 <- datsvy %>% 
  group_by(network1) %>% 
  summarize(DataMean = survey_mean(edge_age_month/12)) %>%
  mutate(DataMean = round(DataMean,3)) %>%
  select(-DataMean_se)

tab <- cbind(means2[,2],means[1])
colnames(tab) <- c("Data Means", "Model Predictions")

datatable(tab)
```

### Transition Networks  

```{r reldur-transition, echo=FALSE}

expFit <- readRDS(here("Survival_Analysis", "model_fits", "covs", "ereltype.rds")) 
cohab_mean <- 1/exp(coef(expFit)[[1]])
spouse_mean <- cohab_mean*(1/exp(coef(expFit)[[2]]))
other_mean <- cohab_mean*(1/exp(coef(expFit)[[3]]))

means <- as.data.frame(rbind(cohab_mean, spouse_mean, other_mean))
means <- round(means/12, 2)
colnames(means) <- c("Model Predictions")
rownames(means) <- c("Cohab", "Spouse", "Other")

dat <- readRDS("~/NSFG_DATA/Objects/alters_active.rds")
datsvy <- as_survey_design(dat, weights = weight, ids=ego)

means2 <- datsvy %>% 
  group_by(reltype) %>% 
  summarize(DataMean = survey_mean(edge_age_month/12)) %>%
  mutate(DataMean = round(DataMean,3)) %>%
  select(-DataMean_se)

tab <- cbind(means2[,2],means[1])
colnames(tab) <- c("Data Means", "Model Predictions")

datatable(tab)
```

## Pseudo-Population Size

pavel recommends: 
"For weighted data the pseudo-populations size should be at least $1 * sampleSize/smallestWeight$ (or $3 * sampleSize/smallestWeight$ to be safe), or 1000 (whichever is larger)."

summary of NSFG weights:
```{r sum-weights}
summary(egodat$egos$weight)
```

NSFG sample size: `r length(egodat$egos$weight)`

```{r ppop-size, echo=FALSE}
ssize <- length(egodat$egos$weight)
minweight <- min(egodat$egos$weight)
```

ppize for $1 * sampleSize/smallestWeight$ = `r round(ssize/minweight)`  
<br>
ppize for $3 * sampleSize/smallestWeight$ = `r round(3*(ssize/minweight))`  

so round up to an even 3500 for estimation? 

when ppop = 3500, you get actual ppop of ~500  
and when ppop = 5000 you get actual pop of ~990 why?  




## Age-Sex-Specific Mortality Rates

raw data pulled from The National Vital Statistics Report [Life Tables for 2017](https://www.cdc.gov/nchs/data/nvsr/nvsr68/nvsr68_07-508.pdf) 

```{r num-survivors, echo=FALSE}
agecat <- c(15, 20, 25, 30, 35, 40)
females <- c(99280, 99134, 98883, 98543, 98083, 97493, 96697)
males <- c(99107, 98749, 98071, 97235, 96284, 95196, 93903)

fdeaths <- c(females[1]-females[2], 
             females[2]-females[3], 
             females[3]-females[4],
             females[4]-females[5],
             females[5]-females[6],
             females[6]-females[7])

mdeaths <- c(males[1]-males[2], 
             males[2]-males[3], 
             males[3]-males[4],
             males[4]-males[5],
             males[5]-males[6],
             males[6]-males[7])

f_death_rate <- round(fdeaths/females[-7], 4)
m_death_rate <- round(mdeaths/males[-7],4)

survivors <- data.frame(row.names=as.character(agecat), females[-7], males[-7], fdeaths, mdeaths, f_death_rate, m_death_rate)
colnames(survivors) <- c("Survs, F", "Survs, M", "Deaths, F", "Deaths, M", "D Rate, F", "D Rate, M")

datatable(survivors, caption="Survivors and Death Rates - 2017")
```

