---
title: "More Data for Duration Sims, YOung"
author: "Emily Pollock"
date: "10/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(plotly)
library(srvyr)
library(reshape2)
library(kableExtra)
library(ergm.ego)
#library(scales)
#library(DataCombine)
#library(DT)
```

More data NSFG for duration ergm sims

Model Params:

- proportion sexually debuted by age, sex  
- time to relationship transition  


Target Stats:  

- mean rel lengths by reltype    
- inst events rate (per year)  



```{r load-data, echo=FALSE}
nsfg <- readRDS("~/NSFG_DATA/Full/nsfg_complete.rds")
nsfg <- nsfg %>% filter(age<30)
svy <- as_survey_design(nsfg, weights = weight, ids = ego)

# young egodata
egodat <- readRDS("~/NSFG_DATA/Objects/fullEgodataYoung.rds")
```

```{r debut, echo=FALSE}
debut <- svy %>% 
          mutate(sex = as.factor(sex), age = as.factor(age), hadsex = as.factor(HADSEX)) %>%
          group_by(sex, age, hadsex) %>% 
          summarize(proportion = survey_mean())

debuted <- debut[which(debut$hadsex==1),] 
debuted <- dcast(debuted, age ~ sex, value.var = "proportion")

debuted %>% plot_ly(x=~age, y=~`F`, name="F", type='scatter', mode="lines") %>%
    add_trace(y=~M, name="M") %>%
    layout(title="Proportion Sexually Debuted", yaxis=list(title="Proportion"), xaxis=list(title="Age"))
```

__Q: at what point do we say everyone is eligible for relationships?__  
(even up to age 44 not 100% sexually debuted)  


## Duration of relationship by type 

mean of active relationships (where we assume exponential dist)
plus
estimates from survival analysis (using all reported relationships), exponential model fit 

```{r reldur, echo=FALSE, include=FALSE}
expFit <- readRDS(here("Survival_Analysis", "model_fits", "covs", "ereltype.rds")) 
cohab_mean <- 1/exp(coef(expFit)[[1]])
spouse_mean <- cohab_mean*(1/exp(coef(expFit)[[2]]))
other_mean <- cohab_mean*(1/exp(coef(expFit)[[3]]))

means <- as.data.frame(rbind(cohab_mean, spouse_mean, other_mean))
means[,2] <- means[,1]/12 
colnames(means) <- c("Months", "Years")

dat <- readRDS("~/NSFG_DATA/Objects/altersegos_survdat.rds")
datsvy <- as_survey_design(dat ,weights = weight, ids=ego)

kable(means)

datsvy %>% filter(active==1) %>% 
  group_by(reltype) %>% 
  summarize(mean_months = survey_mean(edge_age_month),mean_yrs = survey_mean(edge_age_month/12)) %>%
  select(-mean_months_se, -mean_yrs_se)
```

```{r reldur-young, echo=FALSE}
expFityoung <- readRDS(here("Survival_Analysis", "model_fits", "1529", "ereltype.rds")) 

cohab_mean_y <- 1/exp(coef(expFityoung)[[1]])
spouse_mean_y <- cohab_mean*(1/exp(coef(expFityoung)[[2]]))
other_mean_y <- cohab_mean*(1/exp(coef(expFityoung)[[3]]))

means <- as.data.frame(rbind(cohab_mean_y, spouse_mean_y, other_mean_y))
means[,1] <- means[,1]/12 
colnames(means) <- c("Mean Years")

dat2 <- readRDS("~/NSFG_DATA/Objects/alters_young_for_young_egos.rds")
datsvy2 <- as_survey_design(dat2 ,weights = weight, ids=ego)

means2 <-  datsvy2 %>% filter(active==1) %>% group_by(reltype) %>% 
  summarize(mean_yrs = survey_mean(edge_age_month/12))
means2[,1] <- c("Cohab", "Spouse", "Other")

```

### Means
`r kable(means2, format="html")`

### Model Fits
`r kable(means, format="html")`

## Transition Times
other -> cohab (time between rel start and cohabitation)  
cohab -> marriage (time between cohabitation and marriage)  
other -> marriage (time between rel start and marriage, no premarital cohabitation)  

what proportion of all rels are marriages, cohabs, others at cross-section?

### Current Cohabs or Marriages (who cohabitated premaritally)
```{r rt-currentcohabsandmarrs, echo=FALSE}
egosCM <- egodat$egos[which(egodat$egos$ego %in% egodat$altersMarCoh$ego),]

marcohs <- merge(egodat$altersMarCoh, egosCM[,c("ego", "CMINTVW", "cmcohab")], by="ego")

marcohs <- marcohs %>% 
            filter(!is.na(cmcohab)) %>%
            mutate(cmstart = CMINTVW-dfs) %>%
            mutate(timetoCohab = cmcohab-cmstart) %>%
            mutate(preCohabSex = ifelse(timetoCohab>0, "yes", 
                       ifelse(timetoCohab==0, "at cohab",
                       ifelse(timetoCohab <0, "no", NA))))


marcohSvy <- as_survey_design(marcohs, weights = weight, ids=ego)

# other to cohab (either current cohab or current marriage)
otc <- marcohSvy %>% summarize(meanYears = survey_mean(timetoCohab/12, na.rm=T))

# if you limit to only those who report positive
otcp <- marcohSvy %>% filter(timetoCohab > 0) %>%
  summarize(meanYears = survey_mean(timetoCohab/12, na.rm=T))

# if you limit to only those who report negative
otcn <- marcohSvy %>% filter(timetoCohab < 0) %>%
  summarize(meanYears = survey_mean(timetoCohab/12, na.rm=T))

# what prop report cohabitating prior to sex 
preCsex <- marcohSvy %>% 
  filter(!is.na(preCohabSex)) %>%
  group_by(preCohabSex) %>%
  summarize(prop=survey_mean())

#histogram of raw #s
marcohs %>%
  ggplot(aes(x = timetoCohab/12, weights = weight)) +
  geom_histogram(binwidth = 0.5) +
  scale_x_continuous(name="years", limits = c(-10,10)) +
  ggtitle("Time to Cohabitation from Initiation of Sexual Contact")



```

#### Mean Years to Cohabitation      
`r kable(otc, format="html")`  


#### Mean Years to Cohabitation-cohab after sex only    
`r kable(otcp, format="html")`  


#### Mean Years to Cohabitation-cohab before sex only   
`r kable(otcn, format="html")`  


#### Pre-Cohab Sex?    
`r kable(preCsex[,1:2], format="html")`  



### Current Cohabs ONLY 
```{r rt-currentcohabs, echo=FALSE}
egosCohabs <- egodat$egos[which(egodat$egos$ego %in% egodat$altersCohabs$ego),]

cohabs <- merge(egodat$altersCohabs, egosCohabs[,c("ego", "CMINTVW", "cmcohab")], by="ego")
cohabs <- cohabs %>% 
            filter(!is.na(cmcohab)) %>%
            mutate(cmstart = CMINTVW-dfs) %>%
            mutate(timetoCohab = cmcohab-cmstart) %>%
            mutate(preCohabSex = ifelse(timetoCohab>0, "yes", 
                       ifelse(timetoCohab==0, "at cohab",
                       ifelse(timetoCohab <0, "no", NA))))

cohabSvy <- as_survey_design(cohabs, weights = weight, ids=ego)

# other to cohab
otco <- cohabSvy %>% summarize(meanTime = survey_mean(timetoCohab/12, na.rm=T))

# if you limit to only those who report positive
otcop <- cohabSvy %>% filter(timetoCohab > 0) %>%
  summarize(meanTime = survey_mean(timetoCohab/12, na.rm=T))

# if you limit to only those who report negative
otcon <- cohabSvy %>% filter(timetoCohab < 0) %>%
  summarize(meanTime = survey_mean(timetoCohab/12, na.rm=T))

preCsex2 <- cohabSvy %>% 
  filter(!is.na(preCohabSex)) %>%
  group_by(preCohabSex) %>%
  summarize(prop=survey_mean())

#histogram of raw #s
cohabs %>%
  ggplot(aes(x = timetoCohab/12, weights = weight)) +
  geom_histogram(binwidth = 0.5) +
  scale_x_continuous(name="years", limits = c(-10,10)) +
  ggtitle("Time to Cohabitation from Initiation of Sexual Contact")
```

#### Mean Years to Cohabitation  
`r kable(otco, format="html")`   

#### Mean Years to Cohabitation-cohab after sex only    
`r kable(otcop, format="html")`  

#### Mean Years to Cohabitation-cohab before sex only     
`r kable(otcon, format="html")`   

#### Pre-Cohab Sex?    
`r kable(preCsex2[,1:2], format="html")`



### Cohab to Marriage 
```{r rt-cohabtomarraige, echo=FALSE}
egosmarr <- egodat$egos[which(egodat$egos$ego %in% egodat$altersMarriages$ego),]

marrs <- merge(egodat$altersMarriages, egosmarr[,c("ego", "CMINTVW", "cmcohab", "cmmarr")], by="ego")

marrs <- marrs %>% 
  mutate(pmcohab = ifelse(is.na(cmcohab), "no", "yes")) %>%
  mutate(cmstart = CMINTVW-dfs) %>%
  mutate(cohabtoMarr = cmmarr-cmcohab) %>%
  mutate(othertoMarr = cmmarr-cmstart)

marrSvy <- as_survey_design(marrs, weights = weight, ids=ego)

# how many people report cohabitating pre-maritally?
pmc <- marrSvy %>% group_by(pmcohab) %>% summarize(prop = survey_mean())

# cohab to marriage
ctm <- marrSvy %>% summarize(meanYears = survey_mean(cohabtoMarr/12, na.rm=T))
```

#### Pre-Marital Cohabitation  
`r kable(pmc, "html")`

#### Mean time cohab to marriage
`r kable(ctm, "html")`  


```{r marrhist, echo=FALSE}
#histogram of raw #s
marrs %>%
  filter(!is.na(cohabtoMarr)) %>%
  ggplot(aes(x = cohabtoMarr/12, weights = weight)) +
  geom_histogram(binwidth = 0.5) +
  scale_x_continuous(name="years", limits = c(0,10)) +
  ggtitle("Time to Marriage from Initial Cohab")
```



### Other to Marriage
```{r rt-othertomarraige, echo=FALSE}
# other to marriage
otm <- marrSvy %>% summarize(meanYears = survey_mean(othertoMarr/12, na.rm=T))

kable(otm, "html")


#histogram of raw #s
marrs %>%
  ggplot(aes(x = othertoMarr/12, weights = weight)) +
  geom_histogram(binwidth = 0.5) +
  scale_x_continuous(name="years", limits = c(0,10)) +
  ggtitle("Time to Marriage from Initiation of Sexual Contact")
```




## Inst Partners 
average number of inst partners per year per person by sex and age/agecat

### All alters
(mean one-times in last year by sex, race, agecat- inactive one-time partners)
```{r one-times, echo=FALSE}
inst_agecat <- svy %>% 
  filter(age < 30) %>%
  filter(HADSEX==1) %>% 
  group_by(sex, agecat) %>% 
  summarize(mean = survey_mean(instyr)) %>% 
  dcast(sex~agecat, value.var = "mean")

kable(inst_agecat, caption = "By Sex & Ego Race") %>% 
  kable_styling(full_width = F, bootstrap_options = "striped", position="left")


inst_age <- svy %>% 
  filter(age < 30) %>%
  filter(HADSEX==1) %>% 
  group_by(sex, age) %>% 
  summarize(mean = survey_mean(instyr)) %>% 
  mutate(mean = round(mean, 3)) %>%
  dcast(sex~age, value.var = "mean")

kable(inst_age, caption = "By Sex & Ego Race") %>% 
  kable_styling(full_width = F, bootstrap_options = "striped", position="left")
```


### All alters
(mean one-times in last year by sex, race, agecat- inactive one-time partners)
```{r one-times2, echo=FALSE}
inst_agecat <- svy %>% 
  filter(age < 30) %>%
  filter(HADSEX==1) %>% 
  group_by(sex, agecat) %>% 
  summarize(mean = survey_mean(instyr.young)) %>% 
  dcast(sex~agecat, value.var = "mean")


kable(inst_agecat, caption = "By Sex & Ego AgeCat") %>% 
  kable_styling(full_width = F, bootstrap_options = "striped", position="left")


inst_age <- svy %>% 
  filter(age < 30) %>%
  filter(HADSEX==1) %>% 
  group_by(sex, age) %>% 
  summarize(mean = survey_mean(instyr.young)) %>% 
  dcast(sex~age, value.var = "mean")

kable(inst_age, caption = "By Sex & Ego Age") %>% 
  kable_styling(full_width = F, bootstrap_options = "striped", position="left")
```