---
title: "Counterfactual Estimation -- Relationship Duration Models"
author: "Emily Pollock"
date: "11/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# libs
library(ergm.ego)
library(EpiModel)

# data
dat <- readRDS("~/NSFG_DATA/Objects/fullEgodata.rds")

# adjust attributes to what epimodelHIV will expect
# sex = male 1 or 0
dat$egos$male <- ifelse(dat$egos$sex %in% "F", 0, 1)
dat$altersOther$male <- ifelse(dat$altersOther$sex %in% "F", 0, 1)
dat$altersMarCoh$male <- ifelse(dat$altersMarCoh$sex %in% "F", 0, 1)
# debuted = HADSEX
dat$egos$debuted <- ifelse(dat$egos$HADSEX==1, 1, 0)

# agecat, 1:6
breaks <- c(15, 20, 25, 30, 35, 40, 45)
dat$egos$agecat <- cut(dat$egos$age, breaks, labels=F, right=F)
dat$altersOther$agecat <- cut(dat$altersOther$age, breaks, labels=F, right=F)
dat$altersMarCoh$agecat <- cut(dat$altersMarCoh$age, breaks, labels=F, right=F)

#all alters
egodat <- egodata(egos=dat$egos, alters=dat$altersAllActive, egoWt = dat$egos$weight, egoIDcol = "ego")

# marriage/cohab alters
egodat_marcoh <- egodata(egos=dat$egos, alters=dat$altersMarCoh, egoWt = dat$egos$weight, egoIDcol = "ego")

# other alters 
egodat_other <- egodata(egos=dat$egos, alters=dat$altersOther, egoWt = dat$egos$weight, egoIDcol = "ego")
```

# model estimation using ergm.ego

## Marriage/Cohab Network {.tabset .tabset-pills .tabset-fade}

note that these models don't include a condition for degree in other network 

### Fit w/ Nodecov(age) {.tabset .tabset-pills .tabset-fade}

#### Pop = 5,000
```{r marcoh-network}
fit.marcoh.nodecovD <- ergm.ego(egodat_marcoh ~ edges + 
                         nodecov("age") + 
                         absdiff("sqrtage") + 
                         offset(nodematch("male", diff = FALSE)) +
                         offset("concurrent") +
                         offset(nodefactor("debuted", levels=1)), 
                         offset.coef = c(-Inf, -Inf, -Inf),
                         control = control.ergm.ego(ppopsize = 5000))

saveRDS(fit.marcoh.nodecov, "~/Documents/Dissertation/R/Duration/Setup/fits/marcoh-nodecov.RDS")

fit.marcoh.nodecov.5000 <- readRDS("~/Documents/Dissertation/R/Duration/Setup/fits/marcoh-nodecov.RDS")

summary(fit.marcoh.nodecov.5000)

mcmc.diagnostics(fit.marcoh.nodecov.5000)

plot(gof(fit.marcoh.nodecov.5000, GOF = "model"))

plot(gof(fit.marcoh.nodecov.5000, GOF = "degree"))
```

#### Pop = 10,000
```{r marcoh-network-10}
fit.marcoh.nodecov.10 <- ergm.ego(egodat_marcoh ~ edges + 
                         nodecov("age") + 
                         absdiff("sqrtage") + 
                         offset(nodematch("male", diff = FALSE)) +
                         offset(nodefactor("debuted", levels=1)), 
                         offset("concurrent"), 
                         offset.coef = c(-Inf, -Inf),
                         control = control.ergm.ego(ppopsize = 10000))

saveRDS(fit.marcoh.nodecov.10, "~/Documents/Dissertation/R/Duration/Setup/fits/marcoh-nodecov-10.RDS")

fit.marcoh.nodecov.10 <- readRDS("~/Documents/Dissertation/R/Duration/Setup/fits/marcoh-nodecov-10.RDS")

summary(fit.marcoh.nodecov.10)

mcmc.diagnostics(fit.marcoh.nodecov.10)

plot(gof(fit.marcoh.nodecov.10, GOF = "model"))

plot(gof(fit.marcoh.nodecov.10, GOF = "degree"))
```

#### Pop = 20,000
```{r marcoh-network-20}
fit.marcoh.nodecov.20 <- ergm.ego(egodat_marcoh ~ edges + 
                         nodecov("age") + 
                         absdiff("sqrtage") + 
                         offset(nodematch("male", diff = FALSE)) +
                         offset("concurrent"), 
                         offset.coef = c(-Inf, -Inf),
                         control = control.ergm.ego(ppopsize = 20000, ergm.control = control.ergm(MCMLE.maxit = 100)))

saveRDS(fit.marcoh.nodecov.20, "~/Documents/Dissertation/R/Duration/Setup/fits/marcoh-nodecov-20.RDS")

fit.marcoh.nodecov.20 <- readRDS("~/Documents/Dissertation/R/Duration/Setup/fits/marcoh-nodecov-20.RDS")

summary(fit.marcoh.nodecov.20)

mcmc.diagnostics(fit.marcoh.nodecov.20)

plot(gof(fit.marcoh.nodecov.20, GOF = "model"))

plot(gof(fit.marcoh.nodecov.20, GOF = "degree"))
```


### Fit w/ Nodefactor(agecat) {.tabset .tabset-pills .tabset-fade}
#### Pop = 5,000
```{r marcoh-nodefactor-5, echo=FALSE}
fit.marcoh.nodefactor.5 <- ergm.ego(egodat_marcoh ~ edges + 
                         nodefactor("agecat", levels=c(1:3)) +
                         absdiff("sqrtage") + 
                         offset(nodematch("male", diff = FALSE)) +
                         offset("concurrent"), 
                         offset.coef = c(-Inf, -Inf),
                         control = control.ergm.ego(ppopsize = 5000))

saveRDS(fit.marcoh.nodefactor.5, "~/Documents/Dissertation/R/Duration/Setup/fits/marcoh-nodefactor-5.RDS")

fit.marcoh.nodefactor.5 <- readRDS("~/Documents/Dissertation/R/Duration/Setup/fits/marcoh-nodefactor-5.RDS")

summary(fit.marcoh.nodefactor.5)

mcmc.diagnostics(fit.marcoh.nodefactor.5)

plot(gof(fit.marcoh.nodefactor.5, GOF = "model"))

plot(gof(fit.marcoh.nodefactor.5, GOF = "degree"))

```

#### Pop = 10,000
```{r marcoh-nodefactor-10, echo=FALSE}
fit.marcoh.nodefactor.10 <- ergm.ego(egodat_marcoh ~ edges + 
                         nodefactor("agecat", levels=c(1:3)) + 
                         absdiff("sqrtage") + 
                         offset(nodematch("male", diff = FALSE)) +
                         offset("concurrent"), 
                         offset.coef = c(-Inf, -Inf),
                         control = control.ergm.ego(ppopsize = 10000))

saveRDS(fit.marcoh.nodefactor.10, "~/Documents/Dissertation/R/Duration/Setup/fits/marcoh-nodefactor-10.RDS")

fit.marcoh.nodefactor.10 <- readRDS("~/Documents/Dissertation/R/Duration/Setup/fits/marcoh-nodefactor-10.RDS")

summary(fit.marcoh.nodefactor.10)

mcmc.diagnostics(fit.marcoh.nodefactor.10)

plot(gof(fit.marcoh.nodefactor.10, GOF = "model"))

plot(gof(fit.marcoh.nodefactor.10, GOF = "degree"))
```


#### Pop = 20,000
```{r marcoh-nodefactor-20, echo=FALSE}
fit.marcoh.nodefactor.20 <- ergm.ego(egodat_marcoh ~ edges + 
                         nodefactor("agecat", levels=c(1:3)) + 
                         absdiff("sqrtage") + 
                         offset(nodematch("male", diff = FALSE)) +
                         offset("concurrent"), 
                         offset.coef = c(-Inf, -Inf),
                         control = control.ergm.ego(ppopsize = 20000, ergm.control = control.ergm(MCMLE.maxit = 100)))

saveRDS(fit.marcoh.nodefactor.20, "~/Documents/Dissertation/R/Duration/Setup/fits/marcoh-nodefactor-20.RDS")

fit.marcoh.nodefactor.20 <- readRDS("~/Documents/Dissertation/R/Duration/Setup/fits/marcoh-nodefactor-20.RDS")

summary(fit.marcoh.nodefactor.20)

mcmc.diagnostics(fit.marcoh.nodefactor.20)

plot(gof(fit.marcoh.nodefactor.20, GOF = "model"))

plot(gof(fit.marcoh.nodefactor.20, GOF = "degree"))
```

## "Other" formation model:  {.tabset .tabset-pills .tabset-fade}

note these models also don't have contrainst based on degree in marcoh network

### Nodecov(age)
```{r other-network-nodecov, echo=FALSE}
fit.other.cov <- ergm.ego(egodat_other ~ edges + 
                        nodecov("age") + 
                        absdiff("sqrtage") + 
                        concurrent +
                        offset(nodematch("male", diff = FALSE)),
                        offset.coef = c(-Inf),
                        control = control.ergm.ego(ppopsize = 10000))

saveRDS(fit.other.cov, "~/Documents/Dissertation/R/Duration/Setup/fits/other-nodecov.RDS")

fit.other.cov <- readRDS("~/Documents/Dissertation/R/Duration/Setup/fits/other-nodecov.RDS")

summary(fit.other.cov)

mcmc.diagnostics(fit.other.cov)

plot(gof(fit.other.cov, GOF = "model"))

plot(gof(fit.other.cov, GOF = "degree"))
```

### Nodefactor(agecat)
```{r other-network, echo=FALSE}
fit.other <- ergm.ego(egodat_other ~ edges + 
                        nodefactor("agecat", levels=c(1:4)) + 
                        absdiff("sqrtage") + 
                        concurrent + 
                        offset(nodematch("male", diff = FALSE)),
                        offset.coef = c(-Inf),
                        control = control.ergm.ego(ppopsize = 10000))

saveRDS(fit.other, "~/Documents/Dissertation/R/Duration/Setup/fits/other.RDS")

fit.other <- readRDS("~/Documents/Dissertation/R/Duration/Setup/fits/other.RDS")

summary(fit.other)

mcmc.diagnostics(fit.other)

plot(gof(fit.other, GOF = "model"))

plot(gof(fit.other, GOF = "degree"))
```


# NetDX {.tabset .tabset-pills .tabset-fade}
```{r netest-objs, echo=FALSE}
# source JKB's function
source("~/Documents/Dissertation/R/Duration/Setup/ee-netest.R")

# need dissolution stuff

#' # Single dissolution
#' library(EpiModel)
#' diss = ~offset(edges)
#' coef.diss <- dissolution_coefs(dissolution = diss,
#'                                duration = 400,
#'                                d.rate = 3.5e-5)
#'
#' est <- ee.netest(egofit, coef.diss)
#'
#' # Netdx, static
#' dx1 <- netdx(est, nsims = 1e4, dynamic = FALSE,
#'              nwstats.formula = ~edges + meandeg + concurrent)
#' dx1
#' plot(dx1, method = "b", stats = c("edges", "concurrent"))
#'
#' # netdx, dynamic
#' dx2 <- netdx(est, nsims = 5, nsteps = 500,
#'              nwstats.formula = ~edges + meandeg + concurrent,
#'              set.control.ergm = control.simulate.ergm(MCMC.burnin = 1e6))
#' dx2
#' plot(dx2, stats = c("edges", "meandeg"), plots.joined = FALSE)
#' plot(dx2, type = "duration")
#' plot(dx2, type = "dissolution", qnts.col = "orange2")
#' plot(dx2, type = "dissolution", method = "b", col = "bisque")

# marriage dissolution
diss <- ~offset(edges)
# convert from years to weeks
duration <- 8.914 * 52
# death rate
drate <- 3.5e-5

# all together now 
coef.diss <- dissolution_coefs(dissolution=diss,
                               duration=duration,
                               d.rate=drate)
```

## MARCOH 10,000 node netwoks {.tabset .tabset-pills .tabset-fade}
### Nodecov(age)
#### Static 
```{r netdx, echo=FALSE}
marcohnetest <- ee.netest(fit.marcoh.nodecov.10, coef.diss)

# Netdx, static
dx1 <- netdx(marcohnetest, nsims = 1e4, dynamic = FALSE,
  nwstats.formula = ~edges + meandeg + concurrent)

saveRDS(dx1, "~/Documents/Dissertation/R/Duration/Setup/fits/dx-marcohNodecov.RDS")

dx1 <- readRDS("~/Documents/Dissertation/R/Duration/Setup/fits/dx-marcohNodecov.RDS")

dx1

plot(dx1, method = "b", stats = c("edges", "concurrent"))
```

#### Dynamic  
```{r netdx-marcohnodecov-dynamic, echo=FALSE}
# netdx, dynamic
dx2 <- netdx(marcohnetest, nsims = 5, nsteps = 5000,
              nwstats.formula = ~edges + meandeg + concurrent + nodecov("age") + absdiff("sqrtage"),
              set.control.ergm = control.simulate.ergm(MCMC.burnin = 1e6))

saveRDS(dx2, "~/Documents/Dissertation/R/Duration/Setup/fits/dx-marcohNodecov-dynamic.RDS")

dx2 <- readRDS("~/Documents/Dissertation/R/Duration/Setup/fits/dx-marcohNodecov-dynamic.RDS")
               
dx2

plot(dx2, stats = c("edges", "meandeg", "nodecov.age", "absdiff.sqrtage"), plots.joined = FALSE)

plot(dx2, type = "duration")
plot(dx2, type = "dissolution", qnts.col = "orange2")
#plot(dx2, type = "dissolution", method = "b", col = "bisque")
```

### Nodefactor(agecat)
#### Static  
```{r netdx-nodefactormarcoh, echo=FALSE}
marcohnetest2 <- ee.netest(fit.marcoh.nodefactor.10, coef.diss)

# Netdx, static
dx3 <- netdx(marcohnetest2, nsims = 1e4, dynamic = FALSE,
  nwstats.formula = ~edges + meandeg + concurrent)

saveRDS(dx3, "~/Documents/Dissertation/R/Duration/Setup/fits/dx-marcohNodefactor.RDS")

dx3 <- readRDS("~/Documents/Dissertation/R/Duration/Setup/fits/dx-marcohNodefactor.RDS")
               
dx3

plot(dx3, method = "b", stats = c("edges", "concurrent"))
```

#### Dynamic  
```{r netdx-marcohnodefactor-dynamic, echo=FALSE}
# netdx, dynamic
dx4 <- netdx(marcohnetest2, nsims = 5, nsteps = 5000,
              nwstats.formula = ~edges + meandeg + concurrent + nodefactor("agecat", levels=c(1:3)) + absdiff("sqrtage"),
              set.control.ergm = control.simulate.ergm(MCMC.burnin = 1e6))

saveRDS(dx4, "~/Documents/Dissertation/R/Duration/Setup/fits/dx-marcohNodefactor-dynamic.RDS")

dx4 <- readRDS("~/Documents/Dissertation/R/Duration/Setup/fits/dx-marcohNodefactor-dynamic.RDS")
 
dx4

plot(dx4, stats = c("edges", "meandeg", "nodefactor.agecat.1", "nodefactor.agecat.2", "nodefactor.agecat.3", "absdiff.sqrtage"), plots.joined = FALSE)

plot(dx4, type = "duration")
plot(dx4, type = "dissolution", qnts.col = "orange2")
#plot(dx4, type = "dissolution", method = "b", col = "bisque")
```


## OTHER 10,000 node networks {.tabset .tabset-pills .tabset-fade}
### Nodecov(age)
#### Static
```{r other-netdx, echo=FALSE}
# other dissolution
diss <- ~offset(edges)
# convert from years to weeks
duration <- 1.824 * 52
# death rate
drate <- 3.5e-5

# all together now 
coef.diss <- dissolution_coefs(dissolution=diss,
                               duration=duration,
                               d.rate=drate)

otherNetest <- ee.netest(fit.other.cov, coef.diss)

# Netdx, static
dxstatic <- netdx(otherNetest, nsims = 1e4, dynamic = FALSE,
  nwstats.formula = ~edges + meandeg + concurrent)

saveRDS(dxstatic, "~/Documents/Dissertation/R/Duration/Setup/fits/dx-other-cov.RDS")

dxstatic <- readRDS("~/Documents/Dissertation/R/Duration/Setup/fits/dx-other-cov.RDS")
               
dxstatic

plot(dxstatic, method = "b", stats = c("edges","concurrent"))
```

#### Dynamic
```{r netother-dynamic, echo=FALSE}
dx4 <- netdx(otherNetest, nsims = 5, nsteps = 5000,
              nwstats.formula = ~edges + meandeg + concurrent + nodecov("age") + absdiff("sqrtage"),
              set.control.ergm = control.simulate.ergm(MCMC.burnin = 1e6))

saveRDS(dx4, "~/Documents/Dissertation/R/Duration/Setup/fits/dx-other-cov-dynamic.RDS")

dx4 <- readRDS("~/Documents/Dissertation/R/Duration/Setup/fits/dx-other-cov-dynamic.RDS")
 
dx4 

plot(dx4, stats = c("edges", "meandeg", "nodecov.age", "absdiff.sqrtage"), plots.joined = FALSE)

plot(dx4, type = "duration")
plot(dx4, type = "dissolution", qnts.col = "orange2")
```

### Nodefactor(agecat)
#### Static 
```{r other-nodefactor, echo=FALSE}
otherNetest2 <- ee.netest(fit.other, coef.diss)

# Netdx, static
dxstatic2 <- netdx(otherNetest2, nsims = 1e4, dynamic = FALSE,
  nwstats.formula = ~edges + meandeg + concurrent)

saveRDS(dxstatic2, "~/Documents/Dissertation/R/Duration/Setup/fits/dx-other-nodefactor.RDS")

dxstatic2 <- readRDS("~/Documents/Dissertation/R/Duration/Setup/fits/dx-other-nodefactor.RDS")
               
dxstatic2

plot(dxstatic2, method = "b", stats = c("edges","concurrent"))
```
#### Dynamic
```{r other-nodefactor-dynamic, echo=FALSE}
dx_dyn <- netdx(otherNetest2, nsims = 5, nsteps = 5000,
              nwstats.formula = ~edges + meandeg + concurrent + nodefactor("agecat", levels=c(1:3)) + absdiff("sqrtage"),
             set.control.ergm = control.simulate.ergm(MCMC.burnin = 1e6))

saveRDS(dx_dyn, "~/Documents/Dissertation/R/Duration/Setup/fits/dx-other-nodefactor-dynamic.RDS")

dx_dyn <- readRDS("~/Documents/Dissertation/R/Duration/Setup/fits/dx-other-nodefactor-dynamic.RDS")
 
dx_dyn

plot(dx_dyn, stats = c("edges", "meandeg", "nodefactor.agecat.1", "nodefactor.agecat.2", "nodefactor.agecat.3", "absdiff.sqrtage"), plots.joined = FALSE)

plot(dx_dyn, type = "duration")
plot(dx_dyn, type = "dissolution", qnts.col = "orange2")
```

