# CSSS 544 - Final Project

__Question: Assuming no covariate effects, is there evidence of a non-constant hazard for the duration of relationships in the NSFG data?__

__Note:__ Kirk points out that: "the mean of extant partnership ages on the day of interview is considered as an unbiased estimator of the mean of all partnership durations, since both the downwards bias from using censored (incomplete) durations and upwards bias from length bias censoring cancelled out exactly, for proof details, see [Krivitsky et al., 2009]"
 However, because I am not just working with the active partnerships, I have to control for the left truncation and right censoring
 

```{r data}
dat <- readRDS("~/nsfg_analysis/data/alters_egos.rds")

## remove valid skips, missings, once partners
dat <- dat[which(dat$active < 2),]
dat <- dat[which(dat$once==0),]

## new variables

# time
# t_o = relationship dur at 1 year prior to interview = edge_age_month-12, if edge_age_month-12 is negative, 0
# t_c = relationship month end (time = 1-12), if still active, 12
dat$t_o <- ifelse((dat$edge_age_month - 12 >= 0), dat$edge_age_month-12, 0)
dat$t_c <- dat$edge_age_month

# censoring variable (0 if censored, 1 if exact duration data) (active vs inactive)
dat$censored <- ifelse(dat$active==1, 0, 1)

# flag for left truncation correction
dat$truncated <- ifelse(dat$t_o > 0, 1, 0)

# set data up for maxLik package
x <- dat
```
# Example Plot of Known and Unknown Rels based on Sampling Strategy  
```{r example_plot, echo=FALSE}

#### data for plots #### 

# observed rels
examples <- dat[1:75,]
examples <- examples %>% select(dfs, dls, active) %>% mutate(dfs=-dfs, dls=-dls) %>% mutate(dfs_c = dls, dls_c=-dfs)
examples$id <- c(1:nrow(examples))

# left truncated rels 
second <- dat[76:150,] %>% select(dfs, dls, active) %>% mutate(dfs=-dfs, dls=-dls)
offset <- sample(12:30, 75, replace=T)
second <- cbind(second,offset)
second <- second %>% mutate(dfs=dfs-offset, dls=dls-offset)
second <- second[1:37,]
second$id <- seq(1,74, by=2)

# right censored rels 
rights <- examples %>% filter(dls==0 & active==1)
rights$dls_c <- sample(1:80, nrow(rights), replace=T)

#### plot ####
plot(x=NA,
  xlim=c(-350, 60), xaxt='na', xlab=NA,
  ylim=c(0, 75), yaxt='n', ylab=NA,
  main = "Example: Known and Unknown Relationships in NSFG"
)
# obs window
abline(v=-12, lty=2)
abline(v=0, lty=2)
# observed rels
segments(
  x0=examples$dls,
  x1=examples$dfs,
  y0=order(examples$id), 
  lty=1, lwd=2
)
# truncated (unobserved) rels 
segments(
  x0=second$dls,
  x1=second$dfs,
  y0=second$id,
  lty=3, lwd=2
)
# right-censored examples
segments(
  x0=rights$dfs_c,
  x1=rights$dls_c,
  y0=rights$id, 
  lty=2, lwd=2
)
# legend
legend(-348, 17, legend = c("Known Durations", "Possible Right Censored Durations", "Possible Unobserved (Truncated) Durations"), 
       lwd=2, lty=c(1:3), cex=0.6)
```
# Sidebar: what proportion of respondents report more than one relationship? 
```{r serials}

# get ego ids for those who report at least 2 
egos <- dat$ego[which(dat$alter==2)]
length(egos)

tot <- length(unique(dat$ego))

prop <- length(egos) / tot

```
`r prop*100`%


# Modified Kaplan-Meier Estimator  

```{r k-m}
km <- survfit(
  formula = Surv(time = dat$t_o, time2 = dat$t_c, event = dat$censored) ~ 0,
  error = "greenwood"
  )
```


# Model Fitting

## Exponential 

```{r exp-ml}
expML <- function(param){
  lambda <- param
  
  truncated <- x$truncated
  censored <- x$censored
  open <- x$t_o
  closed <- x$t_c

  
  ifelse(truncated== 0,  # if not truncated 
      ifelse(censored == 1,  
             # if exact dur 
             dexp(closed, lambda, log=TRUE),
             #if right censored 
             pexp(closed, lambda, log=TRUE, lower.tail = FALSE)),
      # if truncated
        ifelse(censored == 1, 
               # if exact dur
               dexp(closed, lambda, log=TRUE)/
                 pexp(open, lambda, log=TRUE, lower.tail = FALSE), 
               # if right censored
               pexp(closed, lambda, log=TRUE, lower.tail = FALSE)/
                 pexp(open, lambda, log=TRUE, lower.tail = FALSE))
         )
}

expMaxLik<- maxLik(logLik = expML,  start=c(scale=1), method="BHHH")

summary(expMaxLik)
exp(coef(expMaxLik))
AIC(expMaxLik)


```


```{r mixed-exp}
# exponential mixture model of 2 classes of relationship length
# (similar setup to Model 4 of Kirk's dissertation chapter 2)

# NOTE THAT WE DO NOT USED LOG = TRUE HERE
# we need to add funcations up first and then log them (see parens)


logit = function(p){
  log(p/(1-p))
  }

invLogit = function(phi){
  1/(1+exp(-phi))
  }


mixML <- function(param){
  logLambda1 = param[1]
  logLambda2 = param[2]
  logitWeight1 = param[3]
  
  truncated <- x$truncated
  censored <- x$censored
  open <- x$t_o
  closed <- x$t_c
  
  
  ifelse(truncated == 0,
    ifelse(censored == 1,
            yes = log(
              dexp(closed, exp(logLambda1)) * invLogit(logitWeight1) +
              dexp(closed, exp(logLambda2)) * (1 - invLogit(logitWeight1))
              ),
            no = log(
              pexp(closed, exp(logLambda1), lower.tail = FALSE) * invLogit(logitWeight1) +
              pexp(closed, exp(logLambda2), lower.tail = FALSE) * (1 - invLogit(logitWeight1))
              )
            ),
    ifelse(censored == 1,
      yes = log(
        (dexp(closed, exp(logLambda1)) * invLogit(logitWeight1) +
        dexp(closed, exp(logLambda2)) * (1 - invLogit(logitWeight1))) / 
          
        (pexp(open, exp(logLambda1), lower.tail = FALSE) * invLogit(logitWeight1) +
        pexp(open, exp(logLambda2), lower.tail = FALSE) * (1 - invLogit(logitWeight1)))
        ),
      no = log(
        (pexp(closed, exp(logLambda1), lower.tail = FALSE) * invLogit(logitWeight1) +
        pexp(closed, exp(logLambda2), lower.tail = FALSE) * (1 - invLogit(logitWeight1))) / 
        
          (pexp(open, exp(logLambda1), lower.tail = FALSE) * invLogit(logitWeight1) +
        pexp(open, exp(logLambda2), lower.tail = FALSE) * (1 - invLogit(logitWeight1)))
        )
      )
   )
}


mixExpStartParam = c(
  logLambda1 = log(0.2),
  logLambda2 = log(0.01),
  logitWeight1 = logit(0.5)
  )

mixExpFit = maxLik(
  logLik = mixML,
  start = mixExpStartParam,
  method = "BHHH"
  )

summary(mixExpFit)
exp(coef(mixExpFit)[1:2])
invLogit(coef(mixExpFit)[3])
AIC(mixExpFit)
```

## Weibull 
```{r weib}
weibML <- function(param){
  shape <- param[1]
  scale <- param[2]
  
  truncated <- x$truncated
  censored <- x$censored
  open <- x$t_o
  closed <- x$t_c
  
  ifelse(truncated== 0,  # if not truncated 
      ifelse(censored == 1,  
             # if exact dur 
             dweibull(closed, shape, scale, log=TRUE),
             #if right censored 
             pweibull(closed, shape, scale, log.p=TRUE, lower.tail = FALSE)),
      # if truncated
        ifelse(censored == 1, 
               # if exact dur
               dweibull(closed, shape, scale, log=TRUE)/
                 pweibull(open, shape, scale, log.p=TRUE, lower.tail = FALSE), 
               # if right censored
               pweibull(closed, shape, scale, log.p=TRUE, lower.tail = FALSE)/
                 pweibull(open, shape, scale, log.p=TRUE, lower.tail = FALSE))
         )
}

weibMaxLik <- maxLik(logLik = weibML,  start=c(shape=1, scale=1), method="BHHH")

summary(weibMaxLik)
exp(coef(weibMaxLik))
AIC(weibMaxLik)

```

## Gamma 
```{r gamma}

gammaML <- function(param){
  shape <- param[1]
  scale <- param[2]
  
  truncated <- x$truncated
  censored <- x$censored
  open <- x$t_o
  closed <- x$t_c
  
  ifelse(truncated== 0,  # if not truncated 
      ifelse(censored == 1,  
             # if exact dur 
             dgamma(closed, shape=shape, scale=scale, log=TRUE),
             #if right censored 
             pgamma(closed, shape=shape, scale=scale, log.p=TRUE, lower.tail = FALSE)),
      # if truncated
        ifelse(censored == 1, 
               # if exact dur
               dgamma(closed, shape=shape, scale=scale, log=TRUE)/
                 pgamma(open, shape=shape, scale=scale, log.p=TRUE, lower.tail = FALSE), 
               # if right censored
               pgamma(closed, shape=shape, scale=scale, log.p=TRUE, lower.tail = FALSE)/
                pgamma(open, shape=shape, scale=scale, log.p=TRUE, lower.tail = FALSE))
         )
}

gammaMaxLik <- maxLik(logLik = gammaML,  start=c(shape=1, scale=1), method="BHHH")

summary(gammaMaxLik)
exp(coef(gammaMaxLik))
AIC(gammaMaxLik)

```

```{r plot}

durVec <- 0:max(dat$t_c)

plot(
  km, conf.int = FALSE,
  xlab = "t", ylab = "S(t)"
  )
rug(dat$t_c, lwd=0.07)
lines(x = durVec,
      y = pexp(durVec, coef(expMaxLik), lower.tail = FALSE),
      col="red")
#lines(x= durVec,
#      y=pweibull(durVec, shape=coef(weibMaxLik)[1], scale=coef(weibMaxLik)[2], lower.tail = FALSE), 
#      col="blue")
lines(x= durVec,
      y=pgamma(durVec, shape=coef(gammaMaxLik)[1], scale=coef(gammaMaxLik)[2], lower.tail = FALSE), 
      col="darkgreen")
lines(x = durVec,
      y = invLogit(coef(mixExpFit)[3]) * 
        pexp(q = durVec, exp(coef(mixExpFit)[1]), lower.tail = FALSE) +
        (1-invLogit(coef(mixExpFit)[3])) *
      pexp(q = durVec, exp(coef(mixExpFit)[2]), lower.tail = FALSE),
  col ="purple"
  )

legend(300,1, legend = c("Kaplan-Meier", "Exponential", "Mixed Exponential", "Gamma"),
       col = c("black", "red", "purple", "darkgreen"), cex=0.7, lwd=1)

# black and white plot for poster (and without weibull) 
plot(
  km, conf.int = FALSE,
  xlab = "t", ylab = "S(t)",
  main = "Relationship Survival in the NSFG", lwd=2)
rug(dat$t_c, lwd=0.01)
lines(x = durVec,
      y = pexp(durVec, coef(expMaxLik), lower.tail = FALSE),
      lty=2, lwd=2)
lines(x= durVec,
      y=pgamma(durVec, shape=coef(gammaMaxLik)[1], scale=coef(gammaMaxLik)[2], lower.tail = FALSE), 
      lty=4, lwd=2)
lines(x = durVec,
      y = invLogit(coef(mixExpFit)[3]) * 
        pgeom(q = durVec, prob = exp(coef(mixExpFit)[1]), lower.tail = FALSE) +
        (1-invLogit(coef(mixExpFit)[3])) *
      pgeom(q = durVec, prob = exp(coef(mixExpFit)[2]), lower.tail = FALSE),
  lty=3, lwd=2
  )
legend(300,1, legend = c("Kaplan-Meier", "Exponential", "Mixed Exponential", "Gamma"),
       lty = c(1, 2, 3, 4), lwd=2)

# plot zoomed to first 100 months 

#small_vec <- 0:100

#plot(
#  km, conf.int = FALSE,
#  xlab = "t", ylab = "S(t)",
#  xlim=c(1,100), lwd=2
#  )
#rug(dat$t_c[1:100], lwd=0.07)
#lines(x = small_vec,
#     y = pexp(small_vec, coef(expMaxLik), lower.tail = FALSE),
#      col="red")
#lines(x= small_vec,
#      y=pweibull(small_vec, shape=coef(weibMaxLik)[1], scale=coef(weibMaxLik)[2], lower.tail = FALSE), 
#      col="blue")
#lines(x= small_vec,
#      y=pgamma(small_vec, shape=coef(gammaMaxLik)[1], scale=coef(gammaMaxLik)[2], lower.tail = FALSE), 
#      col="darkgreen")
#lines(x = small_vec,
#      y = invLogit(coef(twoGroupsNexpFit)[3]) * 
#        pgeom(q = small_vec, prob = exp(coef(twoGroupsNexpFit)[1]), lower.tail = FALSE) +
#       (1-invLogit(coef(twoGroupsNexpFit)[3])) *
#      pgeom(q = small_vec, prob = exp(coef(twoGroupsNexpFit)[2]), lower.tail = FALSE),
# col ="purple")

```
```{r aic}



```

# AIC
Exponential: `r AIC(expMaxLik)`  
Gamma: `r AIC(gammaMaxLik)`  
Mixed Exponential: `rAIC(mixExpFit)`  

## Expectations (average rel durs based on each model fit), CIs not calculated
```{r means}
Eexp <- 1/coef(expMaxLik)
Egamma <- coef(gammaMaxLik)[1] * coef(gammaMaxLik)[2]
Emix <- (1/exp(coef(mixExpFit)[1]))*invLogit(coef(mixExpFit)[3]) + 
        (1/exp(coef(mixExpFit)[2]))*(1-invLogit(coef(mixExpFit)[3]))
```

Exponential: `r Eexp` months  
Gamma: `r Egamma` months  
Mixed Exponential:  
- Component 1: `r 1/exp(coef(mixExpFit)[1])` months, `r (invLogit(coef(mixExpFit)[3]))*100`%
- Component 2: `r 1/exp(coef(mixExpFit)[2])` months


__ KNOWN ISSUES__

__K-M CIs:__ hughes et al reccommend a bootstrapping technique or robust variance for CI estimation of left-truncated and right-censored relationship duration data. I do not attempt this here.  

__repeated obvserations per ind__ not dealing with here, `r prop*100`%, and roughly half of those report 3 partners  

__weighting__ no weighting here, can incorporate in to K-M, harder when hand-coding ML for exp and weib, outside of scope of this class project



# things i tried and failed or don't need

```{r mixed-gamma}
# THIS DOESN'T WORK, model either runs for ages or fails

mixMLgamma <- function(param){
  logShape1 = param[1]
  logScale1 = param[2]
  logShape2 = param[3]
  logScale2 = param[4]
  logitWeight1 = param[5]
  
  truncated <- x$truncated
  censored <- x$censored
  open <- x$t_o
  closed <- x$t_c
  
  
  ifelse(truncated == 0,
    ifelse(censored == 1,
      yes = log(
        dgamma(closed, shape=exp(logShape1), scale=exp(logScale1)) * invLogit(logitWeight1) +
        dgamma(closed, shape=exp(logShape2), scale=exp(logScale2)) * (1 - invLogit(logitWeight1))
      ),
      no = log(
        pgamma(closed, shape=exp(logShape1), scale=exp(logScale1), lower.tail = FALSE) * invLogit(logitWeight1) +
        pgamma(closed, shape=exp(logShape2), scale=exp(logScale2), lower.tail = FALSE) * (1 - invLogit(logitWeight1))
      )
    ),
    ifelse(censored == 1,
      yes = log(
        (dgamma(closed, shape=exp(logShape1), scale=exp(logScale1)) * invLogit(logitWeight1) +
        dgamma(closed, shape=exp(logShape2), scale=exp(logScale2)) * (1 - invLogit(logitWeight1))) / 
          
        (pgamma(open, shape=exp(logShape1), scale=exp(logScale1), lower.tail = FALSE) * invLogit(logitWeight1) +
        pgamma(open, shape=exp(logShape2), scale=exp(logScale2), lower.tail = FALSE) * (1 - invLogit(logitWeight1)))
      ),
      no = log(
        (pgamma(closed, shape=exp(logShape1), scale=exp(logScale1), lower.tail = FALSE) * invLogit(logitWeight1) +
        pgamma(closed, shape=exp(logShape2), scale=exp(logScale2), lower.tail = FALSE) * (1 - invLogit(logitWeight1))) / 
          
        (pgamma(open, shape=exp(logShape1), scale=exp(logScale1), lower.tail = FALSE) * invLogit(logitWeight1) +
        pgamma(open, shape=exp(logShape2), scale=exp(logScale2), lower.tail = FALSE) * (1 - invLogit(logitWeight1)))
        )
      )
   )
}


mixGammaStartParam = c(
  logShape1 = log(5),
  logScale1 = log(2),
  logShape2 = log(5),
  logScale2 = log(2),
  logitWeight1 = logit(0.5)
  )

#mixGammaFit = maxLik(
#  logLik = mixMLgamma,
#  start = mixGammaStartParam,
#  method = "BHHH"
#  )

##summary(mixGammaFit)
#exp(coef(mixGammaFit[1:2]))
#invLogit(coef(twoGroupsNexpFit)[3])
#AIC(twoGroupsNexpFit)

```

```{r mixed-gamma_noT}
# this model fit -- why won't the left truncation version????

mixMLgammanoT <- function(param){
  logShape1 = param[1]
  logScale1 = param[2]
  logShape2 = param[3]
  logScale2 = param[4]
  logitWeight1 = param[5]
  
  truncated <- x$truncated
  censored <- x$censored
  open <- x$t_o
  closed <- x$t_c
  
  
    ifelse(censored == 1,
      yes = log(
        dgamma(closed, shape=exp(logShape1), scale=exp(logScale1)) * invLogit(logitWeight1) +
        dgamma(closed, shape=exp(logShape2), scale=exp(logScale2)) * (1 - invLogit(logitWeight1))
      ),
      no = log(
        pgamma(closed, shape=exp(logShape1), scale=exp(logScale1), lower.tail = FALSE) * invLogit(logitWeight1) +
        pgamma(closed, shape=exp(logShape2), scale=exp(logScale2), lower.tail = FALSE) * (1 - invLogit(logitWeight1))
      )
    )
}


mixGammaStartParam = c(
  logShape1 = log(5),
  logScale1 = log(2),
  logShape2 = log(5),
  logScale2 = log(2),
  logitWeight1 = logit(0.5)
  )

#mixGammaFitnoT = maxLik(
#  logLik = mixMLgammanoT,
#  start = mixGammaStartParam,
#  method = "BHHH"
#  )

#summary(mixGammaFitnoT)
#exp(coef(mixGammaFitnoT)[1:4])
#invLogit(coef(mixGammaFitnoT)[5])
#AIC(mixGammaFitnoT)

```

```{r mixed-gamma2}
# this model fit -- why won't the left truncation version????

mixMLgamma2 <- function(param){
  logShape1 = param[1]
  logScale1 = param[2]
  logShape2 = param[3]
  logScale2 = param[4]
  logitWeight1 = param[5]
  
  truncated <- x$truncated
  censored <- x$censored
  open <- x$t_o
  closed <- x$t_c
  
  density <- dgamma(closed, shape=exp(logShape1), scale=exp(logScale1)) * invLogit(logitWeight1) +
        dgamma(closed, shape=exp(logShape2), scale=exp(logScale2)) * (1 - invLogit(logitWeight1))
  
  surv <- pgamma(closed, shape=exp(logShape1), scale=exp(logScale1), lower.tail = FALSE) * invLogit(logitWeight1) +
        pgamma(closed, shape=exp(logShape2), scale=exp(logScale2), lower.tail = FALSE) * (1 - invLogit(logitWeight1))
  
  surv_trunc <- pgamma(open, shape=exp(logShape1), scale=exp(logScale1), lower.tail = FALSE) * invLogit(logitWeight1) +
        pgamma(open, shape=exp(logShape2), scale=exp(logScale2), lower.tail = FALSE) * (1 - invLogit(logitWeight1))
  
  ifelse(truncated==0,  
    ifelse(censored == 1,
      yes = log(density),
      no = log(surv)
    ),
    ifelse(censored == 1,
      yes = log(
        (dgamma(closed, shape=exp(logShape1), scale=exp(logScale1))/pgamma(open, shape=exp(logShape1), scale=exp(logScale1), lower.tail = FALSE)) * invLogit(logitWeight1) +
        (dgamma(closed, shape=exp(logShape2), scale=exp(logScale2))/pgamma(open, shape=exp(logShape2), scale=exp(logScale2), lower.tail = FALSE)) * (1 - invLogit(logitWeight1))
      ),
      
      no = log(
        (pgamma(closed, shape=exp(logShape1), scale=exp(logScale1), lower.tail = FALSE)/pgamma(open, shape=exp(logShape1), scale=exp(logScale1), lower.tail = FALSE)) * invLogit(logitWeight1) +
        (pgamma(closed, shape=exp(logShape2), scale=exp(logScale2), lower.tail = FALSE)/pgamma(open, shape=exp(logShape2), scale=exp(logScale2), lower.tail = FALSE)) * (1 - invLogit(logitWeight1))
      )
    )
  )
}


mixGammaStartParam = c(
  logShape1 = log(5),
  logScale1 = log(2),
  logShape2 = log(5),
  logScale2 = log(2),
  logitWeight1 = logit(0.5)
  )

#mixGammaFit2 = maxLik(
#  logLik = mixMLgamma2,
#  start = mixGammaStartParam,
#  method = "BHHH"
#  )

#summary(mixGammaFit2)
#exp(coef(mixGammaFit2)[1:4])
#invLogit(coef(mixGammaFit2)[5])
#AIC(mixGammaFit2)
```

```{r weib-notrunc}
weibMLnoT <- function(param){
  shape <- param[1]
  scale <- param[2]
  
  censored <- x$censored
  open <- x$t_o
  closed <- x$t_c
  
      ifelse(censored == 1,  
             # if exact dur 
             dweibull(closed, shape, scale, log=TRUE),
             #if right censored 
             pweibull(closed, shape, scale, log.p=TRUE, lower.tail = FALSE))
}

#weibMaxLiknoT <- maxLik(logLik = weibMLnoT,  start=c(shape=1, scale=1), method="BHHH")

#summary(weibMaxLiknoT)

#exp(coef(weibMaxLik))
#AIC(weibMaxLiknoT)

```

```{r mixed-weib}
mixMLweib <- function(param){
  logShape1 = param[1]
  logScale1 = param[2]
  logShape2 = param[3]
  logScale2 = param[4]
  logitWeight1 = param[5]
  
  truncated <- x$truncated
  censored <- x$censored
  open <- x$t_o
  closed <- x$t_c
  
  mixD <- dweibull(closed, exp(logShape1), exp(logScale1)) * invLogit(logitWeight1) +
              dweibull(closed, exp(logShape2), exp(logScale2)) * (1 - invLogit(logitWeight1))
  
  mixS <- pweibull(closed, exp(logShape1), exp(logScale1), lower.tail = FALSE) * invLogit(logitWeight1) +
              pweibull(closed, exp(logShape2), exp(logScale2), lower.tail = FALSE) * (1 - invLogit(logitWeight1))
  
  mixStrunc <- pweibull(open, exp(logShape1), exp(logScale1), lower.tail = FALSE) * invLogit(logitWeight1) +
        pweibull(open, exp(logShape2), exp(logScale2), lower.tail = FALSE) * (1 - invLogit(logitWeight1))
 
 
  ifelse(truncated == 0,
    ifelse(censored == 1,
            yes = log(mixD),
            no = log(mixS)
            ),
    ifelse(censored == 1,
      yes = log(mixD) - log(mixStrunc),
      no = log(mixS) - log(mixStrunc)
      )
   )
}


mixWeibStartParam = c(
  logShape1 = log(0.2), 
  logScale1 = log(0.01),
  logShape2 = log(0.2),
  logScale2 = log(0.01),
  logitWeight1 = logit(0.5)
  )

#mixWeibFit = maxLik(
#  logLik = mixMLweib,
#  start = mixWeibStartParam,
#  method = "BHHH"
#  )

#summary(mixWeibFit)
#exp(coef(mixExpFit)[1:2])
#invLogit(coef(mixExpFit)[3])
#AIC(mixExpFit)

```